---
title: "Gaspereau_2024_Assessment"
output: html_document
---

# Purpose

This markdown file is to have a place where we can:
- prep the data sheets pre-season
- deal with data and do plots in-season for communications
- do analysis and plots for communications post-season

This Rmd file replaces the old "Assessment_Script_RIVER_YEAR.R" scripts

Ideally, this file can be used between years and rivers without much changing, 
where the
year previous to the current one acts as a template. We can use the chunk option
`include=FALSE` to turn a chunk on or off so that we can run the entire file
by clicking the "Run" button at the top of the file in Rstudio.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,    # Hide code
  results = 'hide',# Hide results
  warning = FALSE, # Hide warnings
  message = FALSE  # Hide messages
)
```

```{r library_and_connection}
# Library
library(tidyverse)
library(padr) # for dates
library(ROracle)
```

```{r source_ALOSA_functions}
# Source necessary functions from the Alosa functions
source("~/git/ALOSA.functions/functions/sourcery.R")
sourcery()
```

```{r database_connection}
channel <- dbConnect(
  DBI::dbDriver("Oracle"),
  oracle.username.GASP,
  oracle.password.GASP,
  "PTRAN",
  believeNRows = FALSE
  )
```

# Pre-season

Only run this at the beginning of the season to generate the files we will be
filling out with data throughout the season. Set `include=TRUE` for when you
want to create sheets, other wise it should always be set to FALSE.

```{r create_data_sheets, include=FALSE}

# Creates blank data sheets for counts, biodata, and ages for specified sites
# blank.datasheets(seed=116,startmonth=4,endmonth=6,startday=22,rivername="White Rock",
#                  year=2025,recordtime=T,speciesID=T,strata=5,samplesperstrata=5)
# make.count.filename.textfile("White Rock 2025 count data.csv","WRock",2025)
# 


```

# In-season

run count scripts to get count totals for in season updates
make plots or tables to include in emails

```{r in_season_counts,include=FALSE}
setwd("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2025")

x<-onespecies.river.escapement("White Rock 2025 count data.csv",fixtime=T,database=F,downstream.migration = F,2025,3,channel)

x<-round(x)
# n<-dim(x)[1]
# x<-x[1:n-1,]

print(paste0("Total escapement as of ",x$mon[n],"-",x$day[n]," is ",sum(x$total),sep=""))

# write.csv(x,file="inseasonsummary.csv",row.names=F) # for in season emails
```

```{r in_season_plot,include=FALSE}
#plot for in season emails
old.data<-read.csv("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/data for multi year gaspereau plot.csv")

old.data$date=as.Date(paste(old.data$day,old.data$mon,2025,sep="-"),
                      format="%d-%m-%Y")

#dayofyear uses "strftime" to evaluate which day of the year each date aligns with
old.data$dayofyear=as.numeric(strftime(old.data$date, format="%j"))

plot(x$dayofyear,x$total,type="l", xlim=c(min(x$dayofyear)-4,min(x$dayofyear)+65), ylim=c(0,120000),lwd=2,
     ylab="Number of Fish",xlab="Day of Year")
lines(old.data$dayofyear,old.data$Total2017,type="l",col="brown")
lines(old.data$dayofyear,old.data$Total2018,type="l",col="red")
lines(old.data$dayofyear,old.data$Total2019,type="l",col="orange")
lines(old.data$dayofyear,old.data$Total2021,type="l",col="yellow")
lines(old.data$dayofyear,old.data$Total2022,type="l",col="green")
lines(old.data$dayofyear,old.data$Total2023,type="l",col="blue")
lines(old.data$dayofyear,old.data$Total2024,type="l",col="purple")

legend("topleft",legend=c("2017", "2018", "2019", "2021", "2022", "2023", "2024", "2025"),
       col=c("brown", "red", "orange", "yellow", "green", "blue", "purple", "black"), lty=1, cex=0.8)

```

# Post-Season

clean and prep data for upload
ageing selection
recounts
final versions of some plots

```{r ageing_selection_fishery, include=FALSE}
#first part simulates some data and reads in made up landings data
#second part runs the sampling script on the aggregated data, looping over the three sites
#should be able to swap out the simualted data for real stuff, in the same format,
#to select scales from real data
setwd("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2025")
seed=681:683 #Seed used for scale selection. 
nsamples=300  #Number of scale selected to be aged
sites<-c("LOWER","MIDDLE","UPPER")
#dates set manually for 2025
date.df<-data.frame(fishing.day=1:24,
                    month=c(rep(4,6),rep(5,18)),
                    day=c(21,22,24,25,28,29,01,02,05,06,08,09,12,13,15,16,19,20,22,23,26,27,29,30)
)

real.bio.data<-read.csv("Gaspereau_fishingsites_biocharacterisitcs.csv")
real.bio.data<-merge(real.bio.data,date.df)

daily.landings.all<-read.csv("Gaspereau River 2025 fishery landings.csv")
daily.landings.all<-merge(daily.landings.all,date.df)
daily.landings.all$number.fish<-ifelse(daily.landings.all$unit=="lbs",
                                       daily.landings.all$landings*2,
                                       daily.landings.all$landings*4.4)

###sample it in a for loop for each site
sample.nums<-list()
for(j in 1:3)
{
  #subset to site
  ind<-sites[j] #idk why but it doesn;t work indexing to sites[j]
  daily.landings<-daily.landings.all[daily.landings.all$Location==ind,]
  bio.data<-real.bio.data[real.bio.data$Location==ind,]
  
  n.sampled <- aggregate(bio.data$fishing.day, by = list(bio.data$fishing.day), FUN = function(x){length(x[!is.na(x)])})
  colnames(n.sampled) = c("fishing.day", "n.sampled")
  biodata.with.weights <- merge(bio.data, n.sampled, by = "fishing.day", all.x = T)
  mergedcountdata = aggregate(daily.landings$number.fish, by = list(daily.landings$fishing.day), FUN = sum)
  colnames(mergedcountdata) = c("fishing.day", "merged.number.up")
  biodata.with.weights<-merge(biodata.with.weights, mergedcountdata[,c("fishing.day","merged.number.up")], by="fishing.day", all.x = T)
  
  biodata.with.weights$weighting <- biodata.with.weights$merged.number.up / biodata.with.weights$n.sampled
  # Remove infinite values from the weighting column
  biodata.with.weights$weighting[biodata.with.weights$weighting == Inf] <- 0
  # sample
  set.seed(seed[j])
  ladd.sample <- data.frame(
    sample(biodata.with.weights$sample[biodata.with.weights$scale == 'Y'],
          nsamples,
          replace = F,
          prob = biodata.with.weights$weighting[biodata.with.weights$scale == 'Y']
          )
  )
  colnames(ladd.sample) <- ("sample")
  ladd.sample$Location <- sites[j]
  sample.nums[[j]] <- ladd.sample[with(ladd.sample, order(sample)), ]
}

sample.nums.all<-do.call("rbind",sample.nums)
# write.csv(sample.nums.all,file="GR fishery samples to be aged 2025.csv",row.names=F)

#check for days where all samples are selected (bad)
real.bio.data$to.be.aged<-ifelse(real.bio.data$sample %in% sample.nums.all,1,0)
out<-list()
for(j in 1:3)
{
  ind<-sites[j] #idk why but it doesn;t work indexing to sites[j]
  bio.data<-real.bio.data[real.bio.data$Location==ind,]
  sampled.for.ageing <- aggregate(bio.data$to.be.aged, by = list(bio.data$fishing.day), FUN = mean)
  colnames(sampled.for.ageing) = c("fishing.day", "sampled.for.ageing")
  sampled.for.ageing$site<-sites[j]
  out[[j]]<-sampled.for.ageing
}
sampled.for.ageing<-do.call("rbind",out)
```

```{r final counts}
wr_2025 <- onespecies.river.escapement(
  fixtime = TRUE,
  downstream.migration = FALSE,
  database = TRUE,
  year = 2025,
  site = 3,
  channel = channel
)

#write out counts to include in daily counts all years all rivers

```
```{r fishery and ladder ages}
#get summary data
multiyear=read.csv("//ent.dfo-mpo.ca/atlshares/Science/Population Ecology Division/DFD/Alosa/Assessment Results/river summaries all years.csv",
                   header=T,stringsAsFactors = F)
GR=multiyear[multiyear$river=="Gaspereau" & multiyear$species=="A",]

channel = dbConnect(
  DBI::dbDriver("Oracle"),
  oracle.username.GASP,
  oracle.password.GASP,
  "PTRAN",
  believeNRows = FALSE
) 

wrl25<-get.age.data(2025,siteID = 3,sppID = 3501,AgeStructure = T,PrimaryAger = "Y",channel)

age.prop.ladd.matrix=matrix(rep(0,20),nrow=4, ncol=5,
                       dimnames=list(c("First","Second","Third","Fourth"),
                                     c("3","4","5","6","7")))
for (i in 3:7){
  CurrentAge=wrl25[wrl25$CURRENT_AGE==i,c("CURRENT_AGE","AGE_AT_FIRST_SPAWN")]
  for (j in i:3){
    spawngroup=CurrentAge[CurrentAge$AGE_AT_FIRST_SPAWN==j,]
    proportion=dim(spawngroup)[1]/(dim(wrl25)[1])
    #proportion.weighted=sum(spawngroup$age.weight)/sum(age.dat$age.weight)
    if(proportion>0)age.prop.ladd.matrix[(i-(j-1)),(i-2)]=proportion
    #if(proportion.weighted>0)age.prop.matrix.weighted[(i-(j-1)),(i-2)]=proportion.weighted
  }
}
age.num.ladd.matrix=GR$escape[GR$year==2025 & !is.na(GR$escape)]*age.prop.ladd.matrix/1000

#get fishery age and bio data
grf25<-read.csv("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2025/Ageing/GR fishery samples aged_LILY.csv")
grf25<-grf25[-c(4,7,8,9)]
colnames(grf25)<-c("YEAR","FISH_ID","SITE_ID","CURRENT_AGE","AGE_AT_FIRST_SPAWN")

grfbio25<-read.csv("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2025/Gaspereau_fishingsites_biocharacteristics.csv")
grfbio25<-grfbio25[c(2,3,4,5,7,8,9,11)]
colnames(grfbio25)<-c("YEAR","DAY","MON","FISH_ID","SEX_ID","FORK_LENGTH","WEIGHT","SITE_ID")

#merge ages and bio dataframe to get day and month
grf25<-merge(grf25,grfbio25,by=c("YEAR","FISH_ID","SITE_ID"))

#weight fishery ages
grf.landings25<-read.csv("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2025/Gaspereau River 2025 fishery landings.csv")
colnames(grf.landings25)<-c("YEAR","MON","DAY","SITE_ID","LANDINGS","UNIT")
grf.landings25$LANDINGSFISH<-ifelse(grf.landings25$UNIT=="lbs",grf.landings25$LANDINGS*2,grf.landings25$LANDINGS/0.227)

grf.landings25$DATE<-make_date(year = grf.landings25$YEAR, month = grf.landings25$MON, day = grf.landings25$DAY)
grf.landings25$DOY<-yday(grf.landings25$DATE)

grf25<-merge(grf25,grf.landings25,by=c("YEAR","MON","DAY","SITE_ID"))

grf.numbers.at.age<-function(samp.ages,landings)
{
  grf25<-samp.ages
  grf.landings25<-landings
  #we will scale the aged sample up to the get the total number of aged a,ps fish at each site
  #then scale it back down to the sample size, in numbers-at-age,ps
  
  #treat fishery as single point
  #daily weights
  n.sampled <- aggregate(grf25$DOY, by = list(grf25$DOY), FUN = function(x){length(x[!is.na(x)])})
  #fish caught
  n.caught <- aggregate(grf.landings25$LANDINGSFISH, by = list(grf.landings25$DOY), FUN = "sum")
  weights.df<-merge(n.sampled,n.caught,by="Group.1")
  colnames(weights.df)<-c("DOY","sampled","caught")
  weights.df$age.weight<-weights.df$caught/weights.df$sampled
  
  #merge back with all data to get numbers-at-age
  grf25<-merge(grf25,weights.df[,c(1,4)],by="DOY")
  return(grf25)
}  
grf25.ages<-grf.numbers.at.age(grf25,grf.landings25)


#stolen code
age.prop.fish.matrix=matrix(rep(0,20),nrow=4, ncol=5,
                       dimnames=list(c("First","Second","Third","Fourth"),
                                     c("3","4","5","6","7")))
for (i in 3:7){
  CurrentAge=grf25.ages[grf25.ages$CURRENT_AGE==i,c("CURRENT_AGE","AGE_AT_FIRST_SPAWN","age.weight")]
  for (j in i:3){
    spawngroup=CurrentAge[CurrentAge$AGE_AT_FIRST_SPAWN==j,]
    proportion.weighted=sum(spawngroup$age.weight,na.rm=T)/sum(grf25.ages$age.weight)
    if(proportion.weighted>0)age.prop.fish.matrix[(i-(j-1)),(i-2)]=proportion.weighted
  }
}

age.num.fish.matrix=GR$catch[GR$year==2025 & !is.na(GR$catch)]*age.prop.fish.matrix/1000

#plot
par(mfrow=c(1,2),oma=c(1,0,3,0),xpd=TRUE)
par(mar=c(4.2,5,2,1))
barplot(age.num.ladd.matrix[1:4,],ylim=c(0,800),
        xlab="Age",ylab="Thousands of Fish",cex.lab=1.5,las=1,
        col=c("#E69F00","#56B4E9","#009E73","#0072B2")) 
mtext("White Rock Ladder",3,line=-1.5,las=1,cex=1,adj=0.95)
box()
barplot(age.num.fish.matrix[1:4,],ylim=c(0,800),
        xlab="Age",ylab="Thousands of Fish",cex.lab=1.5,las=1,
        col=c("#E69F00","#56B4E9","#009E73","#0072B2")) 
mtext("Fishery",3,line=-1.5,las=1,cex=1,adj=0.95)
box()
par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
legend("top",legend=c("First","Second","Third","Fourth"),
       fill=c("#E69F00","#56B4E9","#009E73","#0072B2"),bty='n',
       title="Number of Spawnings - 2025",title.adj=0,xpd=TRUE,inset=c(0,0.02),horiz=T)
```
```{r numbers at age all three sites}
#get fishery age and bio data
grf25<-read.csv("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2025/Ageing/GR fishery samples aged_LILY.csv")
grf25<-grf25[-c(4,7,8,9)]
colnames(grf25)<-c("YEAR","FISH_ID","SITE_ID","CURRENT_AGE","AGE_AT_FIRST_SPAWN")

grfbio25<-read.csv("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2025/Gaspereau_fishingsites_biocharacteristics.csv")
grfbio25<-grfbio25[c(2,3,4,5,7,8,9,11)]
colnames(grfbio25)<-c("YEAR","DAY","MON","FISH_ID","SEX_ID","FORK_LENGTH","WEIGHT","SITE_ID")

#merge ages and bio dataframe to get day and month
grf25<-merge(grf25,grfbio25,by=c("YEAR","FISH_ID","SITE_ID"))

#weight fishery ages
grf.landings25<-read.csv("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2025/Gaspereau River 2025 fishery landings.csv")
colnames(grf.landings25)<-c("YEAR","MON","DAY","SITE_ID","LANDINGS","UNIT")
grf.landings25$LANDINGSFISH<-ifelse(grf.landings25$UNIT=="lbs",grf.landings25$LANDINGS*2,grf.landings25$LANDINGS/0.227)

grf.landings25$DATE<-make_date(year = grf.landings25$YEAR, month = grf.landings25$MON, day = grf.landings25$DAY)
grf.landings25$DOY<-yday(grf.landings25$DATE)

grf25<-merge(grf25,grf.landings25,by=c("YEAR","MON","DAY","SITE_ID"))

#separate out sites for later
grf25l<-grf25[grf25$SITE_ID=="LOWER",]
grf25m<-grf25[grf25$SITE_ID=="MIDDLE",]
grf25u<-grf25[grf25$SITE_ID=="UPPER",]

grf.landings25l<-grf.landings25[grf.landings25$SITE_ID=="LOWER",]
grf.landings25m<-grf.landings25[grf.landings25$SITE_ID=="MIDDLE",]
grf.landings25u<-grf.landings25[grf.landings25$SITE_ID=="UPPER",]

grf.numbers.at.age<-function(samp.ages,landings)
{
  grf25<-samp.ages
  grf.landings25<-landings
  #we will scale the aged sample up to the get the total number of aged a,ps fish at each site
  #then scale it back down to the sample size, in numbers-at-age,ps
  
  #treat fishery as single point
  #daily weights
  n.sampled <- aggregate(grf25$DOY, by = list(grf25$DOY), FUN = function(x){length(x[!is.na(x)])})
  #fish caught
  n.caught <- aggregate(grf.landings25$LANDINGSFISH, by = list(grf.landings25$DOY), FUN = "sum")
  weights.df<-merge(n.sampled,n.caught,by="Group.1")
  colnames(weights.df)<-c("DOY","sampled","caught")
  weights.df$age.weight<-weights.df$caught/weights.df$sampled
  
  #merge back with all data to get numbers-at-age
  grf25<-merge(grf25,weights.df[,c(1,4)],by="DOY")
  return(grf25)
} 

grf25.agesl<-grf.numbers.at.age(grf25l,grf.landings25l)
grf25.agesm<-grf.numbers.at.age(grf25m,grf.landings25m)
grf25.agesu<-grf.numbers.at.age(grf25u,grf.landings25u)
grf25.ages<-grf.numbers.at.age(grf25,grf.landings25)


#stolen code
#LOWER
age.prop.low.matrix=matrix(rep(0,20),nrow=4, ncol=5,
                       dimnames=list(c("First","Second","Third","Fourth"),
                                     c("3","4","5","6","7")))
for (i in 3:7){
  CurrentAge=grf25.agesl[grf25.agesl$CURRENT_AGE==i,c("CURRENT_AGE","AGE_AT_FIRST_SPAWN","age.weight")]
  for (j in i:3){
    spawngroup=CurrentAge[CurrentAge$AGE_AT_FIRST_SPAWN==j,]
    proportion.weighted=sum(spawngroup$age.weight,na.rm=T)/sum(grf25.agesl$age.weight)
    if(proportion.weighted>0)age.prop.low.matrix[(i-(j-1)),(i-2)]=proportion.weighted
  }
}

age.num.low.matrix=GR$catch[GR$year==2025 & !is.na(GR$catch)]*age.prop.low.matrix/1000

#MIDDLE
age.prop.mid.matrix=matrix(rep(0,20),nrow=4, ncol=5,
                       dimnames=list(c("First","Second","Third","Fourth"),
                                     c("3","4","5","6","7")))
for (i in 3:7){
  CurrentAge=grf25.agesm[grf25.agesm$CURRENT_AGE==i,c("CURRENT_AGE","AGE_AT_FIRST_SPAWN","age.weight")]
  for (j in i:3){
    spawngroup=CurrentAge[CurrentAge$AGE_AT_FIRST_SPAWN==j,]
    proportion.weighted=sum(spawngroup$age.weight,na.rm=T)/sum(grf25.agesm$age.weight)
    if(proportion.weighted>0)age.prop.mid.matrix[(i-(j-1)),(i-2)]=proportion.weighted
  }
}

age.num.mid.matrix=GR$catch[GR$year==2025 & !is.na(GR$catch)]*age.prop.mid.matrix/1000

#UPPER
age.prop.upp.matrix=matrix(rep(0,20),nrow=4, ncol=5,
                       dimnames=list(c("First","Second","Third","Fourth"),
                                     c("3","4","5","6","7")))
for (i in 3:7){
  CurrentAge=grf25.agesu[grf25.agesu$CURRENT_AGE==i,c("CURRENT_AGE","AGE_AT_FIRST_SPAWN","age.weight")]
  for (j in i:3){
    spawngroup=CurrentAge[CurrentAge$AGE_AT_FIRST_SPAWN==j,]
    proportion.weighted=sum(spawngroup$age.weight,na.rm=T)/sum(grf25.agesu$age.weight)
    if(proportion.weighted>0)age.prop.upp.matrix[(i-(j-1)),(i-2)]=proportion.weighted
  }
}

age.num.upp.matrix=GR$catch[GR$year==2025 & !is.na(GR$catch)]*age.prop.upp.matrix/1000

age.prop.all.matrix=matrix(rep(0,20),nrow=4, ncol=5,
                       dimnames=list(c("First","Second","Third","Fourth"),
                                     c("3","4","5","6","7")))
for (i in 3:7){
  CurrentAge=grf25.ages[grf25.ages$CURRENT_AGE==i,c("CURRENT_AGE","AGE_AT_FIRST_SPAWN","age.weight")]
  for (j in i:3){
    spawngroup=CurrentAge[CurrentAge$AGE_AT_FIRST_SPAWN==j,]
    proportion.weighted=sum(spawngroup$age.weight,na.rm=T)/sum(grf25.ages$age.weight)
    if(proportion.weighted>0)age.prop.all.matrix[(i-(j-1)),(i-2)]=proportion.weighted
  }
}

age.num.all.matrix=GR$catch[GR$year==2025 & !is.na(GR$catch)]*age.prop.all.matrix/1000

#plot
par(mfrow=c(2,2),oma=c(2.5,0.5,4,0),xpd=TRUE)
par(mar=c(2.5,4,0,1))
barplot(age.num.low.matrix[1:4,],ylim=c(0,600),
        xlab="",ylab="Thousands of Fish",cex.lab=1,las=1,
        col=c("#E69F00","#56B4E9","#009E73","#0072B2")) 
mtext("Lower",3,line=-1.5,las=1,cex=1,adj=0.95)
box()
barplot(age.num.mid.matrix[1:4,],ylim=c(0,600),
        xlab="",ylab="",cex.lab=1,las=1,
        col=c("#E69F00","#56B4E9","#009E73","#0072B2")) 
mtext("Middle",3,line=-1.5,las=1,cex=1,adj=0.95)
box()
barplot(age.num.upp.matrix[1:4,],ylim=c(0,600),
        xlab="",ylab="Thousands of Fish",cex.lab=1,las=1,
        col=c("#E69F00","#56B4E9","#009E73","#0072B2")) 
mtext("Upper",3,line=-1.5,las=1,cex=1,adj=0.95)
mtext("Age",1,line=3)
box()
barplot(age.num.all.matrix[1:4,],ylim=c(0,600),
        xlab="",ylab="",cex.lab=1,las=1,
        col=c("#E69F00","#56B4E9","#009E73","#0072B2")) 
mtext("All",3,line=-1.5,las=1,cex=1,adj=0.95)
mtext("Age",1,line=3)
box()
par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
legend("top",legend=c("First","Second","Third","Fourth"),
       fill=c("#E69F00","#56B4E9","#009E73","#0072B2"),bty='n',
       title="Number of Spawnings - 2025",title.adj=0,xpd=TRUE,inset=c(0,0.02),horiz=T)
```


```{r fishery and ladder ages 2016}
#get summary data
multiyear=read.csv("//ent.dfo-mpo.ca/atlshares/Science/Population Ecology Division/DFD/Alosa/Assessment Results/river summaries all years.csv",
                   header=T,stringsAsFactors = F)
GR=multiyear[multiyear$river=="Gaspereau" & multiyear$species=="A",]

channel = dbConnect(
  DBI::dbDriver("Oracle"),
  oracle.username.GASP,
  oracle.password.GASP,
  "PTRAN",
  believeNRows = FALSE
) 

wrl16<-get.age.data(2016,siteID = 3,sppID = 3501,AgeStructure = T,PrimaryAger = "Y",channel)

age.prop.ladd.matrix=matrix(rep(0,20),nrow=4, ncol=5,
                       dimnames=list(c("First","Second","Third","Fourth"),
                                     c("3","4","5","6","7")))
for (i in 3:7){
  CurrentAge=wrl16[wrl16$CURRENT_AGE==i,c("CURRENT_AGE","AGE_AT_FIRST_SPAWN")]
  for (j in i:3){
    spawngroup=CurrentAge[CurrentAge$AGE_AT_FIRST_SPAWN==j,]
    proportion=dim(spawngroup)[1]/(dim(wrl16)[1])
    #proportion.weighted=sum(spawngroup$age.weight)/sum(age.dat$age.weight)
    if(proportion>0)age.prop.ladd.matrix[(i-(j-1)),(i-2)]=proportion
    #if(proportion.weighted>0)age.prop.matrix.weighted[(i-(j-1)),(i-2)]=proportion.weighted
  }
}
age.num.ladd.matrix=GR$escape[GR$year==2016 & !is.na(GR$escape)]*age.prop.ladd.matrix/1000

#get fishery age data
#2016 fishery ages do not require weighting, so treat same as ladder ages
grf16<-read.csv("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2016/Biological characteristics datasheet commercial 05302016 Master withage.csv")
grf16<-grf16[,c(1,2,3,4,6,7,8,11,12)]
colnames(grf16)<-c("DAY","MON","YEAR","FISH_ID","SEX_ID","FORK_LENGTH","WEIGHT","CURRENT_AGE","AGE_AT_FIRST_SPAWN")
grf16<-grf16[complete.cases(grf16[,8]), ] # get rid of trailing NAs
grf16$YEAR<-2016
grf16.ages<-grf16

age.prop.fish.matrix=matrix(rep(0,20),nrow=4, ncol=5,
                       dimnames=list(c("First","Second","Third","Fourth"),
                                     c("3","4","5","6","7")))
for (i in 3:7){
  CurrentAge=grf16.ages[grf16.ages$CURRENT_AGE==i,c("CURRENT_AGE","AGE_AT_FIRST_SPAWN")]
  for (j in i:3){
    spawngroup=CurrentAge[CurrentAge$AGE_AT_FIRST_SPAWN==j,]
    proportion=dim(spawngroup)[1]/(dim(grf16.ages)[1])
    #proportion.weighted=sum(spawngroup$age.weight)/sum(age.dat$age.weight)
    if(proportion>0)age.prop.fish.matrix[(i-(j-1)),(i-2)]=proportion
    #if(proportion.weighted>0)age.prop.matrix.weighted[(i-(j-1)),(i-2)]=proportion.weighted
  }
}

age.num.fish.matrix=GR$catch[GR$year==2016 & !is.na(GR$catch)]*age.prop.fish.matrix/1000

#plot
par(mfrow=c(1,2),oma=c(1,0,3,0),xpd=TRUE)
par(mar=c(4.2,5,2,1))
barplot(age.num.ladd.matrix[1:4,],ylim=c(0,800),
        xlab="Age",ylab="Thousands of Fish",cex.lab=1.5,las=1,
        col=c("#E69F00","#56B4E9","#009E73","#0072B2")) 
mtext("White Rock Ladder",3,line=-1.5,las=1,cex=1,adj=0.95)
box()
barplot(age.num.fish.matrix[1:4,],ylim=c(0,800),
        xlab="Age",ylab="Thousands of Fish",cex.lab=1.5,las=1,
        col=c("#E69F00","#56B4E9","#009E73","#0072B2")) 
mtext("Fishery",3,line=-1.5,las=1,cex=1,adj=0.95)
box()
par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
legend("top",legend=c("First","Second","Third","Fourth"),
       fill=c("#E69F00","#56B4E9","#009E73","#0072B2"),bty='n',
       title="Number of Spawnings - 2016",title.adj=0,xpd=TRUE,inset=c(0,0.02),horiz=T)
```
```{r mean weight}
bio.data<-get.bio.data(2025,3,3501,channel)
count.data<-wr_2025

bio.data$date=as.Date(paste(bio.data$DAY,bio.data$MON,bio.data$YEAR,sep="-"),
                     format="%d-%m-%Y")

#dayofyear uses "strftime" to evaluate which day of the year each date aligns with
bio.data$dayofyear=as.numeric(strftime(bio.data$date, format="%j"))


# scaledata=biodata[biodata$SCALE=="Y",]
n.sampled<-aggregate(bio.data$dayofyear,by=list(bio.data$dayofyear),FUN=function(x){length(x[!is.na(x)])})

colnames(n.sampled)=c("dayofyear","n.sampled")

biodata.with.weights<-merge(bio.data,n.sampled,by="dayofyear",all.x=T)
missingdays<-missing.days(bio.data)
mergedays<-missingdays-1

for(i in 1:length(missingdays)){
  count.data$dayofyear[count.data$dayofyear==missingdays[i]]<-mergedays[i]
}

mergedcountdata=aggregate(count.data$total,by=list(count.data$dayofyear),FUN=sum)
colnames(mergedcountdata)=c("dayofyear","merged.total")

biodata.with.weights<-merge(biodata.with.weights,mergedcountdata[,c("dayofyear","merged.total")],
                            by="dayofyear",all.x=T)


biodata.with.weights$weighting<- biodata.with.weights$merged.total/biodata.with.weights$n.sampled

weighted.mean(biodata.with.weights$WEIGHT,biodata.with.weights$weighting,na.rm=T)

```

