---
title: "Gaspereau_2026_Assessment"
output: html_document
---

# Purpose

This markdown file is to have a place where we can:
- prep the data sheets pre-season
- deal with data and do plots in-season for communications
- do analysis and plots for communications post-season

This Rmd file replaces the old "Assessment_Script_RIVER_YEAR.R" scripts

Ideally, this file can be used between years and rivers without much changing, 
where the
year previous to the current one acts as a template. We can use the chunk option
`include=FALSE` to turn a chunk on or off so that we can run the entire file
by clicking the "Run" button at the top of the file in Rstudio.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,    # Hide code
  results = 'hide',# Hide results
  warning = FALSE, # Hide warnings
  message = FALSE  # Hide messages
)
```

```{r library_and_connection}
# Library
library(tidyverse)
library(padr) # for dates
library(ROracle)
library(CloudBurglar)#for reading in from sharepoint
library(readxl)
```

```{r source_ALOSA_functions}
# Source necessary functions from the Alosa functions
source("~/git/ALOSA.functions/functions/sourcery.R")
sourcery()
```

```{r database_connection}
channel <- dbConnect(
  DBI::dbDriver("Oracle"),
  oracle.username.GASP,
  oracle.password.GASP,
  "PTRAN",
  believeNRows = FALSE
  )
```

# Pre-season

Only run this at the beginning of the season to generate the files we will be
filling out with data throughout the season. Set `include=TRUE` for when you
want to create sheets, other wise it should always be set to FALSE.

```{r create_data_sheets, include=FALSE}

# Creates blank data sheets for counts, biodata, and ages for specified sites
# blank.datasheets(seed=6767,startmonth=4,endmonth=6,startday=15,rivername="White Rock",
#                  year=2026,recordtime=T,speciesID=T,strata=5,samplesperstrata=5)
# make.count.filename.textfile("White Rock 2026 count data.csv","WRock",2026)
# 


```

# In-season

run count scripts to get count totals for in season updates
make plots or tables to include in emails

```{r in_season_counts,include=FALSE}
# setwd("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2026")
counts<-CloudBurglar::get_file("https://086gc.sharepoint.com/:x:/r/sites/msteams_74c888/Shared%20Documents/Alosa/Counts/2026/White%20Rock%202026%20count%20data.xlsx?d=wcf38570fa8344aefb34ae349e4a7b4c5&csf=1&web=1&e=Xo6uvL")


# x<-onespecies.river.escapement("White Rock 2026 count data.csv",fixtime=T,database=F,downstream.migration = F,2026,3,channel)
x<-onespecies.river.escapement(counts,fixtime=T,database=F,downstream.migration = F,2026,3,channel)

x<-round(x)
n<-dim(x)[1]
x<-x[1:n-1,]

print(paste0("Total escapement as of ",x$mon[n],"-",x$day[n]," is ",sum(x$total),sep=""))

# write.csv(x,file="inseasonsummary.csv",row.names=F) # for in season emails
```


```{r in_season_plot, eval=FALSE, include=FALSE}
#plot for in season emails
old.data<-read.csv("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/data for multi year gaspereau plot.csv")

old.data$date=as.Date(paste(old.data$day,old.data$mon,2026,sep="-"),
                      format="%d-%m-%Y")

#dayofyear uses "strftime" to evaluate which day of the year each date aligns with
old.data$dayofyear=as.numeric(strftime(old.data$date, format="%j"))

plot(x$dayofyear,x$total,type="l", xlim=c(min(x$dayofyear)-4,min(x$dayofyear)+65), ylim=c(0,120000),lwd=2,
     ylab="Number of Fish",xlab="Day of Year")
lines(old.data$dayofyear,old.data$Total2017,type="l",col="brown")
lines(old.data$dayofyear,old.data$Total2018,type="l",col="red")
lines(old.data$dayofyear,old.data$Total2019,type="l",col="orange")
lines(old.data$dayofyear,old.data$Total2021,type="l",col="yellow")
lines(old.data$dayofyear,old.data$Total2022,type="l",col="green")
lines(old.data$dayofyear,old.data$Total2023,type="l",col="blue")
lines(old.data$dayofyear,old.data$Total2024,type="l",col="purple")
lines(old.data$dayofyear,old.data$Total2025,type="l",col="violet")

legend("topleft",legend=c("2017", "2018", "2019", "2021", "2022", "2023", "2024", "2025","2026"),
       col=c("brown", "red", "orange", "yellow", "green", "blue", "purple", "violet", "black"), lty=1, cex=0.8)


#cumulative sum plot
old.data[is.na(old.data)]<-0 #to allow cumsum to work

plot(x$dayofyear,cumsum(x$total),type="l", xlim=c(min(x$dayofyear)-4,min(x$dayofyear)+65), ylim=c(0,1400000),lwd=2,
     ylab="Number of Fish",xlab="Day of Year")
lines(old.data$dayofyear,cumsum(old.data$Total2017),type="l",col="brown")
lines(old.data$dayofyear,cumsum(old.data$Total2018),type="l",col="red")
lines(old.data$dayofyear,cumsum(old.data$Total2019),type="l",col="orange")
lines(old.data$dayofyear,cumsum(old.data$Total2021),type="l",col="yellow")
lines(old.data$dayofyear,cumsum(old.data$Total2022),type="l",col="green")
lines(old.data$dayofyear,cumsum(old.data$Total2023),type="l",col="blue")
lines(old.data$dayofyear,cumsum(old.data$Total2024),type="l",col="purple")
lines(old.data$dayofyear,cumsum(old.data$Total2025),type="l",col="violet")

legend("topleft",legend=c("2017", "2018", "2019", "2021", "2022", "2023", "2024", "2025","2026"),
       col=c("brown", "red", "orange", "yellow", "green", "blue", "purple", "violet", "black"), lty=1, cex=0.8)

```

# Post-Season

clean and prep data for upload
ageing selection
recounts
final versions of some plots

```{r ageing_selection, include=FALSE}
setwd("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2026")
seed=546 #Seed used for scale selection. 
nsamples=500  #Number of scale selected to be aged

bio.data<-read.csv("White Rock 2026 biocharacteristics data.csv")

#Check Bio data
# checkers("White Rock 2026 biocharacteristics data.csv") # this file has the correct column names, same data as WE 2023 bio data.csv

#combine extra columns into notes
bio.data$notes<-paste(bio.data$samplers,bio.data$notes,sep=" ")
bio.data$notes<-ifelse(bio.data$otolith=="y",paste0(bio.data$notes," otolith taken",sep=""),bio.data$notes)


#need to do some testing that the age weighting is OK

countdata<-daily.count<-x

bio.data<-bio.data[,c(1:10,13)]
# colnames(bio.data)<-c("SITE_ID", "YEAR", "DAY", "MON","FISH_ID", "SPECIES_ID",
#                                        "SEX_ID","FORK_LENGTH","WEIGHT","SCALE","NOTES") 
biodata<-bio.data

#no missing days for 2026, doing weekly weighting instead
# missingdays<-missing.days(bio.data)
# mergedays<-missingdays-1 #in both cases the two previous days are most similar in count
# For missing sample days, we merge the counts from two days
# and use that in the weighting calculation. 
# For example, ff day 112 is missing then decide if you want to merge the counts
# with day 111 or 113. Do this for all the missing dates and provide the
# replacement days in this vector. Length(mergedays)==Length(missingdays)


#ran July 08th 2026 to get ages for scale mounting
# ageing.selection.test(countdata,bio.data,weekly=TRUE,2026,seed=seed,nsamples=nsamples,species="A")


```

```{r prep data for upload}
#clean up BIodata as above
setwd("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2026")
bio.data<-read.csv("White Rock 2026 biocharacteristics data.csv")

#Check Bio data
checkers("White Rock 2026 biocharacteristics data.csv") #used to check for data entry errors, then corrections can be made to csv

#combine extra columns into notes
bio.data$notes<-paste(bio.data$samplers,bio.data$notes,sep=" ")
bio.data$notes<-ifelse(bio.data$otolith=="y",paste0(bio.data$notes," otolith taken",sep=""),bio.data$notes)

#file can be uploaded with additional columns, only those mathcing database will be uploaded
#go to ALOSA.functions/Writing Annual Data to Oracle.R to upload


```


```{r flow data}
setwd("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2026")
flows26<-read_xlsx("Gaspereau River Apr-June 2026.xlsx",skip=4,col_names=c("TS","Water Level","Estimated Discharge m3s","Estimated Discharge cfs")) # skip the 4 lines to only read in the data. Manually code column names

flows26<-flows26[-c(1:10),] #remove the first 10 data points per Gabriel's email

par(mar=c(5,4,4,4)+0.1)
plot(flows26$TS,flows26$`Estimated Discharge cfs`,type="l",
     main="Gaspereau River Flows and Water Levels 2026",xlab="",ylab="",
     yaxt="n",col="orange")
axis(2,las=2) #for flow in cfs
par(new=T)
plot(flows26$TS,flows26$`Water Level`,type="l",,xlab="",ylab="",
     yaxt="n",col="green")
axis(4,las=2) #for flow in cfs
legend("topright",col=c("orange","green"),legend=c("Estimated Discharge (cfs)","Water Level (m)"),lty=c(1,1))
mtext("Date",1,line=3,cex=1.2)
mtext("Estimated Discharge (cfs)",2,line=3,cex=1.2)
mtext("Water Level (m)",4,line=3,cex=1.2)


#plot 2025 too
setwd("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2025")
flows25<-read_xlsx("Gaspereau River Apr-June 2025 flows.xlsx",skip=1,col_names=c("TS","Water Level","Estimated Discharge m3s","Estimated Discharge cfs"),col_types = c("date","numeric","numeric","numeric")) #Manually code column names and skip first row. specify column types to avoid issues

flows25<-flows25[flows25$TS>"2025-04-13 00:00:00",] #trim to the day the river was lowered to install gear

par(mar=c(5,4,4,4)+0.1)
plot(flows25$TS,flows25$`Estimated Discharge cfs`,type="l",
     main="Gaspereau River Flows and Water Levels 2025",xlab="",ylab="",
     yaxt="n",col="orange")
axis(2,las=2) #for flow in cfs
par(new=T)
plot(flows25$TS,flows25$`Water Level`,type="l",,xlab="",ylab="",
     yaxt="n",col="green")
axis(4,las=2) #for flow in cfs
legend("topright",col=c("orange","green"),legend=c("Estimated Discharge (cfs)","Water Level (m)"),lty=c(1,1))
mtext("Date",1,line=3,cex=1.2)
mtext("Estimated Discharge (cfs)",2,line=3,cex=1.2)
mtext("Water Level (m)",4,line=3,cex=1.2)

#look at flows and counts
#setup count dates
x$date<-as.Date(paste(x$day,x$mon,2026,sep="-"),format="%d-%m-%Y")

#setup x axis limits with dates
xlimdatetimes<-as_datetime(c("2026-04-13 00:00:00","2026-06-23 00:00:00"))
xlimdates<-as.Date(c("2026-04-13","2026-06-23"))
ylims=c(0,950) #max for 26 is 917, standard is 4% extra, 950 pretty close

par(mar=c(5,4,4,4)+0.1)
plot(flows26$TS,flows26$`Estimated Discharge cfs`,type="l",xlim=xlimdatetimes,ylim=ylims,
     main="Gaspereau River Flows and Counts 2026",xlab="",ylab="",
     yaxt="n",col="orange")
axis(2,las=2) #for flow in cfs
par(new=T)
plot(x$date,x$total/1000,type="l",,xlab="",ylab="",xlim=xlimdates,
     yaxt="n",col="black")
axis(4,las=2) #for flow in cfs
legend("topright",col=c("orange","black"),legend=c("Estimated Discharge (cfs)","Counts"),lty=c(1,1))
mtext("Date",1,line=3,cex=1.2)
mtext("Estimated Discharge (cfs)",2,line=3,cex=1.2)
mtext("Thousands of fish",4,line=3,cex=1.2)

#2025 counts
# wr_2025 <- onespecies.river.escapement(
#   fixtime = TRUE,
#   downstream.migration = FALSE,
#   database = TRUE,
#   year = 2025,
#   site = 3,
#   channel = channel
# )
wr_2025$date<-as.Date(paste(wr_2025$day,wr_2025$mon,2025,sep="-"),format="%d-%m-%Y")

xlimdatetimes<-as_datetime(c("2025-04-13 00:00:00","2025-06-18 00:00:00"))
xlimdates<-as.Date(c("2025-04-13","2025-06-18"))

par(mar=c(5,4,4,4)+0.1)
plot(flows25$TS,flows25$`Estimated Discharge cfs`,type="l",xlim=xlimdatetimes,ylim=ylims,
     main="Gaspereau River Flows and Counts 2025",xlab="",ylab="",
     yaxt="n",col="orange")
axis(2,las=2) #for flow in cfs
par(new=T)
plot(wr_2025$date,wr_2025$total/1000,type="l",,xlab="",ylab="",xlim=xlimdates,
     yaxt="n",col="black")
axis(4,las=2) #for flow in cfs
legend("topright",col=c("orange","black"),legend=c("Estimated Discharge (cfs)","Counts"),lty=c(1,1))
mtext("Date",1,line=3,cex=1.2)
mtext("Estimated Discharge (cfs)",2,line=3,cex=1.2)
mtext("Thousands of fish",4,line=3,cex=1.2)

# look at flows on same plot
# need to adjust dates, uses lubridate
year(flows25$TS)<-2026

plot(flows26$TS,flows26$`Estimated Discharge cfs`,type="l",
     main="Gaspereau River Flows 2025 and 2026",xlab="",ylab="",
     yaxt="n",col="orange")
lines(flows25$TS,flows25$`Estimated Discharge cfs`,col="red")
axis(2,las=2) #for flow in cfs
mtext("Date",1,line=3,cex=1.2)
mtext("Estimated Discharge (cfs)",2,line=3,cex=1.2)
legend("topright",col=c("orange","red"),legend=c("2026","2025"),lty=c(1,1))
```

