---
title: "Gaspereau_2026_Assessment"
output: html_document
---

# Purpose

This markdown file is to have a place where we can:
- prep the data sheets pre-season
- deal with data and do plots in-season for communications
- do analysis and plots for communications post-season

This Rmd file replaces the old "Assessment_Script_RIVER_YEAR.R" scripts

Ideally, this file can be used between years and rivers without much changing, 
where the
year previous to the current one acts as a template. We can use the chunk option
`include=FALSE` to turn a chunk on or off so that we can run the entire file
by clicking the "Run" button at the top of the file in Rstudio.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,    # Hide code
  results = 'hide',# Hide results
  warning = FALSE, # Hide warnings
  message = FALSE  # Hide messages
)
```

```{r library_and_connection}
# Library
library(tidyverse)
library(padr) # for dates
library(ROracle)
library(CloudBurglar)#for reading in from sharepoint
```

```{r source_ALOSA_functions}
# Source necessary functions from the Alosa functions
source("~/git/ALOSA.functions/functions/sourcery.R")
sourcery()
```

```{r database_connection}
channel <- dbConnect(
  DBI::dbDriver("Oracle"),
  oracle.username.GASP,
  oracle.password.GASP,
  "PTRAN",
  believeNRows = FALSE
  )
```

# Pre-season

Only run this at the beginning of the season to generate the files we will be
filling out with data throughout the season. Set `include=TRUE` for when you
want to create sheets, other wise it should always be set to FALSE.

```{r create_data_sheets, include=FALSE}

# Creates blank data sheets for counts, biodata, and ages for specified sites
# blank.datasheets(seed=6767,startmonth=4,endmonth=6,startday=15,rivername="White Rock",
#                  year=2026,recordtime=T,speciesID=T,strata=5,samplesperstrata=5)
# make.count.filename.textfile("White Rock 2026 count data.csv","WRock",2026)
# 


```

# In-season

run count scripts to get count totals for in season updates
make plots or tables to include in emails

```{r in_season_counts,include=FALSE}
# setwd("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2026")
counts<-CloudBurglar::get_file("https://086gc.sharepoint.com/:x:/r/sites/msteams_74c888/Shared%20Documents/Alosa/Counts/2026/White%20Rock%202026%20count%20data.xlsx?d=wcf38570fa8344aefb34ae349e4a7b4c5&csf=1&web=1&e=XQ0fC6")

# x<-onespecies.river.escapement("White Rock 2026 count data.csv",fixtime=T,database=F,downstream.migration = F,2026,3,channel)
x<-onespecies.river.escapement(counts,fixtime=T,database=F,downstream.migration = F,2026,3,channel)

x<-round(x)
n<-dim(x)[1]
x<-x[1:n-1,]

print(paste0("Total escapement as of ",x$mon[n],"-",x$day[n]," is ",sum(x$total),sep=""))

write.csv(x,file="inseasonsummary.csv",row.names=F) # for in season emails
```

```{r}
# remotes::install_github("PopulationEcologyDivision/CloudBurglar")
vcount<-CloudBurglar::get_file("https://086gc.sharepoint.com/:x:/r/sites/msteams_74c888/Shared%20Documents/Alosa/Counts/2026/Vaughan%202026%20count%20data.xlsx?d=wbaf213b4e3a04b3aa3abdb1f2c847933&csf=1&web=1&e=C1ehyS")
y<-onespecies.partial.river.escapement(vcount,fixtime=T,database=F,2026,2,channel)

v.strata<-y$summary.data
v.total<-y$daily.summary
v.strata<-v.strata[order(v.strata$dayofyear),]


```



```{r in_season_plot,include=FALSE}
#plot for in season emails
old.data<-read.csv("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/data for multi year gaspereau plot.csv")

old.data$date=as.Date(paste(old.data$day,old.data$mon,2026,sep="-"),
                      format="%d-%m-%Y")

#dayofyear uses "strftime" to evaluate which day of the year each date aligns with
old.data$dayofyear=as.numeric(strftime(old.data$date, format="%j"))

plot(x$dayofyear,x$total,type="l", xlim=c(min(x$dayofyear)-4,min(x$dayofyear)+65), ylim=c(0,120000),lwd=2,
     ylab="Number of Fish",xlab="Day of Year")
lines(old.data$dayofyear,old.data$Total2017,type="l",col="brown")
lines(old.data$dayofyear,old.data$Total2018,type="l",col="red")
lines(old.data$dayofyear,old.data$Total2019,type="l",col="orange")
lines(old.data$dayofyear,old.data$Total2021,type="l",col="yellow")
lines(old.data$dayofyear,old.data$Total2022,type="l",col="green")
lines(old.data$dayofyear,old.data$Total2023,type="l",col="blue")
lines(old.data$dayofyear,old.data$Total2024,type="l",col="purple")
lines(old.data$dayofyear,old.data$Total2025,type="l",col="violet")

legend("topleft",legend=c("2017", "2018", "2019", "2021", "2022", "2023", "2024", "2025","2026"),
       col=c("brown", "red", "orange", "yellow", "green", "blue", "purple", "violet", "black"), lty=1, cex=0.8)


#cumulative sum plot
old.data[is.na(old.data)]<-0 #to allow cumsum to work

plot(x$dayofyear,cumsum(x$total),type="l", xlim=c(min(x$dayofyear)-4,min(x$dayofyear)+65), ylim=c(0,1400000),lwd=2,
     ylab="Number of Fish",xlab="Day of Year")
lines(old.data$dayofyear,cumsum(old.data$Total2017),type="l",col="brown")
lines(old.data$dayofyear,cumsum(old.data$Total2018),type="l",col="red")
lines(old.data$dayofyear,cumsum(old.data$Total2019),type="l",col="orange")
lines(old.data$dayofyear,cumsum(old.data$Total2021),type="l",col="yellow")
lines(old.data$dayofyear,cumsum(old.data$Total2022),type="l",col="green")
lines(old.data$dayofyear,cumsum(old.data$Total2023),type="l",col="blue")
lines(old.data$dayofyear,cumsum(old.data$Total2024),type="l",col="purple")
lines(old.data$dayofyear,cumsum(old.data$Total2025),type="l",col="violet")

legend("topleft",legend=c("2017", "2018", "2019", "2021", "2022", "2023", "2024", "2025","2026"),
       col=c("brown", "red", "orange", "yellow", "green", "blue", "purple", "violet", "black"), lty=1, cex=0.8)

```

# Post-Season

clean and prep data for upload
ageing selection
recounts
final versions of some plots


