---
title: "Gaspereau_2024_Assessment"
output: html_document
---

# Purpose

This markdown file is to have a place where we can:
- prep the data sheets pre-season
- deal with data and do plots in-season for communications
- do analysis and plots for communications post-season

This Rmd file replaces the old "Assessment_Script_RIVER_YEAR.R" scripts

Ideally, this file can be used between years and rivers without much changing, 
where the
year previous to the current one acts as a template. We can use the chunk option
`include=FALSE` to turn a chunk on or off so that we can run the entire file
by clicking the "Run" button at the top of the file in Rstudio.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,    # Hide code
  results = 'hide',# Hide results
  warning = FALSE, # Hide warnings
  message = FALSE  # Hide messages
)
```

```{r library_and_connection}
# Library
library(tidyverse)
library(padr) # for dates
library(ROracle)
```

```{r source_ALOSA_functions}
# Source necessary functions from the Alosa functions
source("~/git/ALOSA.functions/functions/sourcery.R")
sourcery()
```

```{r database_connection}
channel <- dbConnect(
  DBI::dbDriver("Oracle"),
  oracle.username.GASP,
  oracle.password.GASP,
  "PTRAN",
  believeNRows = FALSE
  )
```

# Pre-season

Only run this at the beginning of the season to generate the files we will be
filling out with data throughout the season. Set `include=TRUE` for when you
want to create sheets, other wise it should always be set to FALSE.

```{r create_data_sheets, include=FALSE}

# Creates blank data sheets for counts, biodata, and ages for specified sites
# blank.datasheets(seed=116,startmonth=4,endmonth=6,startday=22,rivername="White Rock",
#                  year=2025,recordtime=T,speciesID=T,strata=5,samplesperstrata=5)
# make.count.filename.textfile("White Rock 2025 count data.csv","WRock",2025)
# 


```

# In-season

run count scripts to get count totals for in season updates
make plots or tables to include in emails

```{r in_season_counts,include=FALSE}
setwd("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River")

x<-onespecies.river.escapement("White Rock 2025 count data.csv",fixtime=T,database=F,downstream.migration = F,2025,3,channel)

x<-round(x)
n<-dim(x)[1]
x<-x[1:n-1,]

print(paste0("Total escapement as of ",x$mon[n],"-",x$day[n]," is ",sum(x$total),sep=""))

write.csv(x,file="inseasonsummary.csv",row.names=F) # for in season emails
```

```{r in_season_plot,include=FALSE}
#plot for in season emails
old.data<-read.csv("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/data for multi year gaspereau plot.csv")

old.data$date=as.Date(paste(old.data$day,old.data$mon,2025,sep="-"),
                      format="%d-%m-%Y")

#dayofyear uses "strftime" to evaluate which day of the year each date aligns with
old.data$dayofyear=as.numeric(strftime(old.data$date, format="%j"))

plot(x$dayofyear,x$total,type="l", xlim=c(min(x$dayofyear)-4,min(x$dayofyear)+65), ylim=c(0,120000),lwd=2,
     ylab="Number of Fish",xlab="Day of Year")
lines(old.data$dayofyear,old.data$Total2017,type="l",col="brown")
lines(old.data$dayofyear,old.data$Total2018,type="l",col="red")
lines(old.data$dayofyear,old.data$Total2019,type="l",col="orange")
lines(old.data$dayofyear,old.data$Total2021,type="l",col="yellow")
lines(old.data$dayofyear,old.data$Total2022,type="l",col="green")
lines(old.data$dayofyear,old.data$Total2023,type="l",col="blue")
lines(old.data$dayofyear,old.data$Total2024,type="l",col="purple")

legend("topleft",legend=c("2017", "2018", "2019", "2021", "2022", "2023", "2024", "2025"),
       col=c("brown", "red", "orange", "yellow", "green", "blue", "purple", "black"), lty=1, cex=0.8)

```

# Post-Season

clean and prep data for upload
ageing selection
recounts
final versions of some plots

```{r ageing_selection_fishery, include=FALSE}
#first part simulates some data and reads in made up landings data
#second part runs the sampling script on the aggregated data, looping over the three sites
#should be able to swap out the simualted data for real stuff, in the same format,
#to select scales from real data
setwd("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2025")
seed=681:683 #Seed used for scale selection. 
nsamples=300  #Number of scale selected to be aged
sites<-c("LOWER","MIDDLE","UPPER")
#dates set manually for 2025
date.df<-data.frame(fishing.day=1:24,
                    month=c(rep(4,6),rep(5,18)),
                    day=c(21,22,24,25,28,29,01,02,05,06,08,09,12,13,15,16,19,20,22,23,26,27,29,30)
)

real.bio.data<-read.csv("Gaspereau_fishingsites_biocharacterisitcs.csv")
real.bio.data<-merge(real.bio.data,date.df)

daily.landings.all<-read.csv("Gaspereau River 2025 fishery landings.csv")
daily.landings.all<-merge(daily.landings.all,date.df)
daily.landings.all$number.fish<-ifelse(daily.landings.all$unit=="lbs",
                                       daily.landings.all$landings*2,
                                       daily.landings.all$landings*4.4)

###sample it in a for loop for each site
sample.nums<-list()
for(j in 1:3)
{
  #subset to site
  ind<-sites[j] #idk why but it doesn;t work indexing to sites[j]
  daily.landings<-daily.landings.all[daily.landings.all$Location==ind,]
  bio.data<-real.bio.data[real.bio.data$Location==ind,]
  
  n.sampled <- aggregate(bio.data$fishing.day, by = list(bio.data$fishing.day), FUN = function(x){length(x[!is.na(x)])})
  colnames(n.sampled) = c("fishing.day", "n.sampled")
  biodata.with.weights <- merge(bio.data, n.sampled, by = "fishing.day", all.x = T)
  mergedcountdata = aggregate(daily.landings$number.fish, by = list(daily.landings$fishing.day), FUN = sum)
  colnames(mergedcountdata) = c("fishing.day", "merged.number.up")
  biodata.with.weights<-merge(biodata.with.weights, mergedcountdata[,c("fishing.day","merged.number.up")], by="fishing.day", all.x = T)
  
  biodata.with.weights$weighting <- biodata.with.weights$merged.number.up / biodata.with.weights$n.sampled
  # Remove infinite values from the weighting column
  biodata.with.weights$weighting[biodata.with.weights$weighting == Inf] <- 0
  # sample
  set.seed(seed[j])
  ladd.sample <- data.frame(
    sample(biodata.with.weights$sample[biodata.with.weights$scale == 'Y'],
          nsamples,
          replace = F,
          prob = biodata.with.weights$weighting[biodata.with.weights$scale == 'Y']
          )
  )
  colnames(ladd.sample) <- ("sample")
  ladd.sample$Location <- sites[j]
  sample.nums[[j]] <- ladd.sample[with(ladd.sample, order(sample)), ]
}

sample.nums.all<-do.call("rbind",sample.nums)
# write.csv(sample.nums.all,file="GR fishery samples to be aged 2025.csv",row.names=F)

#check for days where all samples are selected (bad)
real.bio.data$to.be.aged<-ifelse(real.bio.data$sample %in% sample.nums.all,1,0)
out<-list()
for(j in 1:3)
{
  ind<-sites[j] #idk why but it doesn;t work indexing to sites[j]
  bio.data<-real.bio.data[real.bio.data$Location==ind,]
  sampled.for.ageing <- aggregate(bio.data$to.be.aged, by = list(bio.data$fishing.day), FUN = mean)
  colnames(sampled.for.ageing) = c("fishing.day", "sampled.for.ageing")
  sampled.for.ageing$site<-sites[j]
  out[[j]]<-sampled.for.ageing
}
sampled.for.ageing<-do.call("rbind",out)
```

```{r ageing_selection, include=FALSE}
setwd("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2024")
seed=667 #Seed used for scale selection. 
nsamples=500  #Number of scale selected to be aged

bio.data<-read.csv("White Rock 2024 biocharacteristics data.csv")

#Check Bio data
checkers("White Rock 2024 biocharacteristics data.csv") # this file has the correct column names, same data as WE 2023 bio data.csv

countdata<-daily.count<-x

bio.data<-bio.data[,1:11]
# colnames(bio.data)<-c("SITE_ID", "YEAR", "DAY", "MON","FISH_ID", "SPECIES_ID",
#                                        "SEX_ID","FORK_LENGTH","WEIGHT","SCALE","NOTES") 
biodata<-bio.data
# missingdays<-missing.days(bio.data)
# mergedays<-missingdays-1 #in both cases the two previous days are most similar in count
# For missing sample days, we merge the counts from two days
# and use that in the weighting calculation. 
# For example, ff day 112 is missing then decide if you want to merge the counts
# with day 111 or 113. Do this for all the missing dates and provide the
# replacement days in this vector. Length(mergedays)==Length(missingdays)


#ran June 24th 2024 to get ages for student work      
# ageing.selection(daily.count,bio.data,weekly=FALSE,2024,missingdays,mergedays,seed,nsamples,"A")
# scale.age<-read.csv("to be aged_rename this file.csv")

```

```{r setup_WR_recounts, include=FALSE}
setwd("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2024")
#sim code
set.seed(seed)
counts<-read.csv("White Rock 2024 count data.csv")
counts$number.up<-counts$count.upstream-counts$count.downstream
counts$dev<-rnorm(nrow(counts),1,0.1)
counts$recount<-round(counts$number.up*counts$dev,0)
counts.sub<-counts[counts$number.up!=0,]
counts.sub<-counts.sub[complete.cases(counts.sub),]
counts.sample<-sample(1:nrow(counts.sub),200,replace=F) #this will draw 200 random counts to redo
counts.sub<-counts.sub[counts.sample,]
counts.sub$per.diff<-(counts.sub$number.up-counts.sub$recount)/((counts.sub$number.up+counts.sub$recount)/2)

#we can use this to make a count sheet to redo these randomly selected counts
counts.sub1<-counts.sub[,c(1:5,9)]
names(counts.sub1)[6]<-"previous counter"
counts.sub1<-counts.sub1[order(counts.sub1$mon,counts.sub1$day,counts.sub1$time),]
counts.sub1$count.upstream<-NA
counts.sub1$count.downstream<-NA
counts.sub1$number.up<-NA
counts.sub1$camera.desc<-NA
counts.sub1$minutes<-NA
counts.sub1$seconds<-NA

# write.csv(counts.sub1,"White Rock 2024 RECOUNTS.csv",row.names=F,na="")
```

```{r analyze_WR_recounts,include=FALSE}
setwd("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Gaspereau River/Gaspereau 2024")
#recounts are msotly complete, lets take a look
wr.recounts<-read.csv("White Rock 2024 RECOUNTS.csv")
counts<-read.csv("White Rock 2024 count data.csv")

wr.recounts<-merge(wr.recounts,counts,by=c("year","day","mon","time","strata"))
#get only desired columns for tidiness
wr.recounts<-wr.recounts[,c(1:10,14:17)]
colnames(wr.recounts)<-c("year","day","mon","time","strata","previous.counter","count.upstream","count.downstream","number.up",  "new.counter","og.up","og.down","og.number.up","notes")

#remove outlier
wr.recounts<-wr.recounts[-78,]

wr.recounts$quot<-wr.recounts$count.upstream/wr.recounts$og.up
wr.recounts$per.diff<-(wr.recounts$count.upstream-wr.recounts$og.up)/wr.recounts$og.up
NSP.bias<-sum(wr.recounts$count.upstream,na.rm=T)/sum(wr.recounts$og.up,na.rm=T)

plot(wr.recounts$per.diff)
plot(wr.recounts$count.upstream,wr.recounts$per.diff)
segments(0,0.05,max(wr.recounts$count.upstream,na.rm=T)*1.2,0.05,lty=3)
segments(0,-0.05,max(wr.recounts$count.upstream,na.rm=T)*1.2,-0.05,lty=3)

plot(wr.recounts$og.up,wr.recounts$count.upstream)

```

```{r plot_counts}
countdata<-x
countdata$date<-make_date(2024, countdata$mon, countdata$day)
plot(countdata$date,countdata$chigh,type="l",lty=3,xlab="",ylab="")
lines(countdata$date,countdata$clow,lty=3)
lines(countdata$date,countdata$total,lty=1)
segments(as.Date("2024-05-22"),0,as.Date("2024-05-22"),max(countdata$chigh,na.rm=T))
mtext("Date",1,line=3)
mtext("Number of Fish",2,line=3)
```
```{r post_season_counts}
count.data<-onespecies.river.escapement(fixtime = T,
                                        downstream.migration = F,
                                        year = 2024,
                                        database = T,
                                        site = 3,
                                        channel = channel
                                        )
#yay it works
count.data<-round(count.data,0)
write.csv(count.data,file="Gaspereau River 2024 daily counts.csv",row.names=F)
```
```{r mean weight}
#use count data from above
bio.data<-get.bio.data(2024,3,3501,channel)

bio.data$date=as.Date(paste(bio.data$DAY,bio.data$MON,bio.data$YEAR,sep="-"),
                     format="%d-%m-%Y")

#dayofyear uses "strftime" to evaluate which day of the year each date aligns with
bio.data$dayofyear=as.numeric(strftime(bio.data$date, format="%j"))


# scaledata=biodata[biodata$SCALE=="Y",]
n.sampled<-aggregate(bio.data$dayofyear,by=list(bio.data$dayofyear),FUN=function(x){length(x[!is.na(x)])})

colnames(n.sampled)=c("dayofyear","n.sampled")

biodata.with.weights<-merge(bio.data,n.sampled,by="dayofyear",all.x=T)
missingdays<-missing.days(bio.data)
mergedays<-missingdays-1

for(i in 1:length(missingdays)){
  count.data$dayofyear[count.data$dayofyear==missingdays[i]]<-mergedays[i]
}

mergedcountdata=aggregate(count.data$total,by=list(count.data$dayofyear),FUN=sum)
colnames(mergedcountdata)=c("dayofyear","merged.total")

biodata.with.weights<-merge(biodata.with.weights,mergedcountdata[,c("dayofyear","merged.total")],
                            by="dayofyear",all.x=T)


biodata.with.weights$weighting<- biodata.with.weights$merged.total/biodata.with.weights$n.sampled

weighted.mean(biodata.with.weights$WEIGHT,biodata.with.weights$weighting)

```

