---
title: "`r paste('Annual gaspereau escapement assessment for Tusket River for', format(Sys.Date(), '%Y'))`"
output: html_document
---

# Purpose

This markdown file is to have a place where we can:
- prep the data sheets pre-season
- deal with data and do plots in-season for communications
- do analysis and plots for communications post-season

Ideally, this file can be used between years without much changing, where the
year previous to the current one acts as a template. We can use the chunk option
`include=FALSE` to turn a chunk on or off so that we can run the entire file
by clicking the "Run" button at the top of the file in Rstudio.

# Set-up
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,    # Hide code
  results = 'hide',# Hide results
  warning = FALSE, # Hide warnings
  message = FALSE  # Hide messages
)
```

```{r library_and_connection}
# Library
library(tidyverse)
library(padr) # for dates
library(ROracle)
library(imputeTS)

# Constants
#current_year <- format(Sys.Date(), "%Y")
current_year <- 2025

```

```{r source_ALOSA_functions}
# Source necessary functions from the Alosa functions
source("~/git/ALOSA.functions/functions/sourcery.R")
sourcery()
```

```{r database_connection}
channel <- dbConnect(
  DBI::dbDriver("Oracle"),
  username = "GASPEREA",
  password = "gps983",
  "PTRAN",
  believeNRows = FALSE
  )
```

# Pre-season

Only run this at the beginning of the season to generate the files we will be
filling out with data throughout the season. Set `include=TRUE` for when you
want to create sheets, other wise it should always be set to FALSE.

```{r create_data_sheets, include=FALSE}

# Creates blank data sheets for counts and biodata for Lake Vaughan
# blank.datasheets(
#   seed = 112,
#   startmonth = 3,
#   endmonth = 6,
#   startday = 1,
#   rivername = "Vaughan",
#   year = 2023,
#   recordtime = T,
#   speciesID = T,
#   strata = 6,
#   samplesperstrata = 4
# )
# 
# # Creates blank data sheets for counts and biodata for Powerhouse
# blank.datasheets(
#   seed = 113,
#   startmonth = 3,
#   endmonth = 6,
#   startday = 15,
#   rivername = "Powerhouse",
#   year = 2023,
#   recordtime = T,
#   speciesID = T,
#   strata = 6,
#   samplesperstrata = 4
# )

```

# In-season

These are the counts of gaspereau in real time that we can use for internal
communications. We may want to remove the last day if the counts for that day
are not completed in the counts file, but I have left that out for now.

## VD daily escapement
```{r escapement_vd}
# Locate your CSV for counts
path_to_counts_CSV <- "C:/Users/graylo/Documents/Vaughan 2025 count data.csv"

# Run escapement script
# Note: arguments year, site, and channel don't matter when database = FALSE
escapement_vd <- onespecies.river.escapement(
  path_to_counts_CSV,
  fixtime = TRUE,
  downstream.migration = FALSE,
  database = FALSE,
  year = current_year,
  site = 2,
  channel = channel
)

# Compare this year to last year
# escapement_vd <- onespecies.river.escapement(
#   path_to_counts_CSV,
#   fixtime = TRUE,
#   downstream.migration = FALSE,
#   database = TRUE,
#   year = 2024,
#   site = 2,
#   channel = channel
# )

# We want to add where the data are from
escapement_vd$location <- "Lake Vaughan"

# Write these data to CSV so we can share them in emails
# output_folder <- "your/output/folder/goes/here"
# 
# write.csv(
#   escapement_vd,
#   file = paste0(output_folder, "/vd_in_season_summary.csv"),
#   row.names = FALSE
# )

vd <- escapement_vd %>%
    mutate(across(everything(), ~ ifelse(is.nan(.), 0, .))) %>%
    mutate(across(everything(), ~ ifelse(is.na(.), 0, .)))

vd <- vd |> mutate(date = make_date(current_year, mon, day))

#in_season_vd_escapement_plot <-
vd |>
  ggplot(aes(date, total)) +
  geom_path(
    colour = "#7ECDBB",
    alpha = 0.9,
    linewidth = 1.25
    ) +
  geom_ribbon(
    aes(ymin = clow, ymax = chigh),
    fill = "#7ECDBB",
    colour = "grey90",
    alpha = 0.2
  ) +
  theme_bw() +
  labs(
    title = paste0("Gaspereau escapement at the Lake Vaughan fish ladder in ", current_year),
    x = "Date",
    y = "Fish / day"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "7 day") +
  scale_y_continuous(
    limits = c(0, max(vd$chigh)),
    breaks = seq(0, max(vd$chigh), by = 1e4),
    labels = scales::comma
  ) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      vjust = 1,
      hjust = 1
      ),
    axis.title.x = element_blank(),
    panel.grid.minor = element_blank(),
    #legend.background = element_blank(),
    plot.background = element_rect(fill = "grey95")
    )

# plot_path <- paste0(output_folder, "/in_season_vd_escapement_plot.png")
# ggsave(plot_path, plot = in_season_vd_escapement_plot, width = 6, height = 4, dpi = 300)
#sum(vd$total)
```

## PH daily escapement
```{r escapement_ph}
# Locate your CSV for counts
path_to_counts_CSV <- "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2024/Data Sheets/Powerhouse 2024 count data.csv"

# Run escapement script
# Note: arguments year, site, and channel don't matter when database = FALSE
escapement_ph <- onespecies.river.escapement(
  path_to_counts_CSV,
  fixtime = TRUE,
  downstream.migration = FALSE,
  database = FALSE,
  year = current_year,
  site = 14,
  channel = channel
)

# We want to add where the data are from
escapement_ph$location <- "Powerhouse"

# Write these data to CSV so we can share them in emails
# output_folder <- "your/output/folder/goes/here"
# 
# write.csv(
#   escapement_ph,
#   file = paste0(output_folder, "/ph_in_season_summary.csv"),
#   row.names = FALSE
# )

ph <- escapement_ph %>%
    mutate(across(everything(), ~ ifelse(is.nan(.), 0, .))) %>%
    mutate(across(everything(), ~ ifelse(is.na(.), 0, .)))

ph <- ph |> mutate(date = make_date(2024, mon, day))

#in_season_ph_escapement_plot <-
ph |>
  ggplot(aes(date, total)) +
  geom_path(
    colour = "#7ECDBB",
    alpha = 0.9,
    linewidth = 1.25
    ) +
  geom_ribbon(
    aes(ymin = clow, ymax = chigh),
    fill = "#7ECDBB",
    colour = "grey90",
    alpha = 0.2
  ) +
  theme_bw() +
  labs(
    title = paste0("Gaspereau escapement for Powerhouse for 2024"),
    x = "Date",
    y = "Fish / day"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "2 days") +
  scale_y_continuous(
    limits = c(0, max(ph$chigh)),
    breaks = seq(0, max(ph$chigh), by = 1000),
    labels = scales::comma
  ) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      vjust = 1,
      hjust = 1),
    axis.title.x = element_blank(),
    panel.grid.minor = element_blank(),
    #legend.background = element_blank(),
    plot.background = element_rect(fill = "grey95")
  )

# plot_path <- paste0(output_folder, "/in_season_ph_escapement_plot.png")
#     
#     ggsave(
#       plot_path,
#       plot = in_season_ph_escapement_plot,
#       width = 6,
#       height = 4,
#       dpi = 300
#     )
```

## VD multi-year daily
```{r compare_vd}
# Set up a function to haul data from the db for VD
get_VD_multiple_years <- function() {
  
  counts_df <- data.frame()
  
  # You can change how many years you want here
  get_years <- seq(as.integer(current_year) - 5, as.integer(current_year) - 1)
  
  for (i in get_years) {
    
    tryCatch({
      
      year_data <- onespecies.river.escapement(
        filename = "dummy_file_name",
        fixtime = TRUE,
        downstream.migration = TRUE,
        database = TRUE,
        year = i,
        site = 2,
        channel = channel
      )
      
      year_data$year <- i
      
      counts_df <- rbind(counts_df, year_data)
      
    }, error = function(e) {
      
      return(counts_df)
      
    })
    
  }
  
  return(counts_df)
  
}

multi_year_counts <- get_VD_multiple_years()

vd$year <- current_year
vd <- vd |> select(-location, -date)

multi_year_counts <- rbind(multi_year_counts, vd)

# Make sure to filter before creating dates because this screws them up ¯\_(ツ)_/¯
multi_year_counts <- multi_year_counts %>%
  mutate(across(everything(), ~ ifelse(is.nan(.), 0, .))) %>%
  mutate(across(everything(), ~ ifelse(is.na(.), 0, .)))

# You make these all have the same year in their date so they plot nicely
multi_year_counts <- multi_year_counts |>
  mutate(date = make_date(as.character(current_year), mon, day))

multi_year_counts$year <- as.character(multi_year_counts$year)

# I use the %>% pipe here to use "." to filter data ¯\_(ツ)_/¯
multi_year_counts %>%
  group_by(year) %>%
  ggplot(aes(date, total)) +
  geom_path(
    aes(colour = year),
    alpha = if_else(multi_year_counts$year != current_year, 0.5, 0.9),
    linewidth = if_else(multi_year_counts$year != current_year, 1, 2),
    na.rm = TRUE
  )+
  #scale_colour_viridis_d(direction = -1)+
  theme_bw() +
  labs(
    title = "Gaspereau escapement at Lake Vaughan Dam",
    x = "Date",
    y = "Fish / day",
    colour = "Year",
    fill = "Year",
    linetype = "Location"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  scale_y_continuous(
    limits = c(0, 120000),
    breaks = seq(0, 120000, by = 10000),
    labels = scales::comma
  ) + 
  theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.background = element_blank(),
    plot.background = element_rect(fill = "grey95"),
    legend.key = element_blank()
  ) +
  guides(colour = guide_legend(override.aes = list(linewidth = 2)))

# plot_path <-
#   paste0(output_folder, "/in_season_multi_year_counts_plot.png")
# 
# ggsave(
#   plot_path,
#   plot = in_season_multi_year_counts_plot,
#   width = 6,
#   height = 4,
#   dpi = 300
# )
```

## VD multi-year cumulative
```{r}
# Multi-year cumulative counts

multi_year_cum <- multi_year_counts %>%
  group_by(year) %>%
  arrange(date) %>%
  mutate(cumulative_total = cumsum(total))

multi_year_cum %>%
  group_by(year) %>%
  ggplot(aes(date, cumulative_total)) +
  geom_path(
    aes(colour = year),
    alpha = 0.9,
    linewidth = 1.75,
    na.rm = TRUE
  )+
  #scale_colour_viridis_d(direction = -1)+
  theme_bw() +
  labs(
    title = "Cumulative escapement of gaspereau at Lake Vaughan Dam",
    x = "Date",
    y = "Total fish escaped",
    colour = "Year",
    fill = "Year",
    linetype = "Location"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  scale_y_continuous(
    limits = c(0, 2.5e6),
    breaks = seq(0, 2.5e6, 0.2e6),
    labels = scales::comma
  ) + 
  theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.background = element_blank(),
    plot.background = element_rect(fill = "grey95"),
    legend.key = element_blank()
  ) +
  guides(colour = guide_legend(override.aes = list(linewidth = 2)))
```

## VD vs PH
```{r vd_vs_ph}

vd$location <- "Lake Vaughan"
vd$date <- make_date(vd$year, vd$mon, vd$day)
ph$year <- current_year

counts <- rbind(vd, ph)

counts |>
  ggplot(aes(date, total)) +
  geom_path(
    data = counts |> filter(location == "Lake Vaughan"),
    aes(colour = location),
    alpha = 0.9,
    linewidth = 1.25
  ) +
  geom_ribbon(
    data = counts |> filter(location == "Lake Vaughan"),
    aes(ymin = clow,
        ymax = chigh,
        fill = location,
        colour = location),
    alpha = 0.2
  ) +
  geom_path(
    data = counts |> filter(location == "Powerhouse"),
    aes(colour = location),
    alpha = 0.9,
    linewidth = 1.25
  ) +
  geom_ribbon(
    data = counts |> filter(location == "Powerhouse"),
    aes(ymin = clow,
        ymax = chigh,
        fill = location,
        colour = location),
    alpha = 0.5
  ) +
  theme_bw() +
  labs(
    title = "Escapement estimates for gaspereau on the Tusket River",
    x = "Date",
    y = "fish / day",
    colour = "Location",
    fill = "Location"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  scale_y_continuous(
    limits = c(0, max(counts$chigh)),
    breaks = seq(0, max(counts$chigh), by = 10000),
    labels = scales::comma
  ) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      vjust = 1,
      hjust = 1
    ),
    axis.title.x = element_blank(),
    legend.position = "right",
    legend.background = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "grey95")
    )

# plot_path <- paste0(output_folder, "/in_season_ph_vd_escapement_plot.png")
# 
# ggsave(
#   plot_path,
#   plot = in_season_ph_vd_escapement_plot,
#   width = 6,
#   height = 4,
#   dpi = 300
# )

```

## PH multi-year
```{r multi_year_ph}
ph_counts_2022 <- onespecies.partial.river.escapement(
  "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2022/Data Sheets/Counts/Powerhouse Count Sheet Cleaned 2022.csv",
  fixtime = FALSE,
  database = FALSE,
  2022,
  14,
  channel
  )
ph_counts_2022$year <- as.character(2022)
ph_counts_2022$location <- "Powerhouse"

ph_counts_2023 <- onespecies.river.escapement(
  "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2023/Data Sheets/Powerhouse 2023 count data1.csv",
  fixtime = FALSE,
  downstream.migration = FALSE,
  database = FALSE,
  2023,
  14,
  channel
  )
ph_counts_2023$year <- as.character(2023)
ph_counts_2023$location <- "Powerhouse"

# Calculate the powerhouse estimate
# Locate your CSV for counts
path_to_counts_CSV <- "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2024/Data Sheets/Powerhouse 2024 count data.csv"

# Run escapement script
# Note: arguments year, site, and channel don't matter when database = FALSE
escapement_ph <- onespecies.river.escapement(
  path_to_counts_CSV,
  fixtime = TRUE,
  downstream.migration = FALSE,
  database = FALSE,
  year = 2024,
  site = 14,
  channel = channel
)
escapement_ph$location <- "Powerhouse"
escapement_ph$year <- current_year

ph <- escapement_ph %>%
    mutate(across(everything(), ~ ifelse(is.nan(.), 0, .))) %>%
    mutate(across(everything(), ~ ifelse(is.na(.), 0, .)))

# Combine with counts dataframe that was previously created and which should
# have the PH counts from the in season year. This can be probably be done a
# better way, but where's the fun in that?
counts_for_multi_ph <- rbind(ph, ph_counts_2022, ph_counts_2023)
counts_for_multi_ph <- counts_for_multi_ph |> select(-location)

# Make sure to filter before creating dates because this screws them up ¯\_(ツ)_/¯
counts_for_multi_ph <- counts_for_multi_ph %>%
  mutate(across(everything(), ~ ifelse(is.nan(.), 0, .))) %>%
  mutate(across(everything(), ~ ifelse(is.na(.), 0, .)))

# You make these all have the same year in their date so they plot nicely
counts_for_multi_ph <- counts_for_multi_ph |> mutate(date = make_date(2024, mon, day))
counts_for_multi_ph$year <- as.character(counts_for_multi_ph$year)

counts_for_multi_ph %>%
  group_by(year) %>% 
  ggplot(aes(date, total)) +
  geom_path(
    aes(colour = year),
    alpha = if_else(counts_for_multi_ph$year != current_year, 0.3, 0.8),
    linewidth = if_else(counts_for_multi_ph$year != current_year, 1, 1.2),
    na.rm = TRUE
  )+
  geom_ribbon(
    data = . %>% filter(year == 2024),
    aes(ymin = clow, ymax = chigh),
    fill = "#7eadfc",
    colour = "#7eadfc",
    alpha = 0.2
  )+
  theme_bw() +
  labs(
    title = "Gaspereau escapement for the Powerhouse",
    x = "Date",
    y = "Fish / day",
    colour = "Year"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "3 days") +
  scale_y_continuous(
    limits = c(0, max(counts_for_multi_ph$chigh)),
    breaks = seq(0, max(counts_for_multi_ph$chigh), by = 1000),
    labels = scales::comma
  ) +
  theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.background = element_rect(fill = "grey95"),
    legend.background = element_blank(),
    legend.key = element_blank()
  )

```

# Post-season

The analyses here can only be done after uploading the data from the year to the
database. Read the `Writing Annual Data to Oracle.R` script in the `Alosa.functions`
folder to see the protocol for uploading data. There are methods in that script
to prevent errors, check the data, etc.

## Cleaning / uploading biodata
```{r}
# Use the checkers function to check the biocharacteristics data for outliers
# and mistakes
checkers(filename = "R:/Science\Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2025/data/Vaughan 2025 biocharacteristics data.csv")

# Once you go through the checkers function, you can add the accessory data
# (i.e. the additional species ID checks we do for bluebacks) to the biodata.
# The accessory data have to be done a certain way for this to work. The data 
# need to be categorized as BB identified and all fish identified. This may
# be different than you record the data throughout the season, so make sure that
# your data are represented properly:
# Column names: mon, day, BB.sampled, all.sampled, BB.ID, all.ID
accessory <- read_csv(file = "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2025/data/vaughan-accessory-data-version-2.csv")

# WARNING: this writes to the "biofilename" file, it appends the data to the
# end of the file. If you run this more than one it will keep appending the
# data to the file. NOTE: make sure the order of the columns for the date in 
# the biocharacteristics CSV are as follows: day, mon, year
speciesID.expand(
  data = accessory,
  # The biocharacteristics data file path goes here
  biofilename = "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2025/data/Vaughan 2025 biocharacteristics data.csv",
  rivername = "Tusket",
  year = 2025
)

# Here are the new biodata
biodata <- read_csv(file = "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2025/data/Vaughan 2025 biocharacteristics data.csv")

# Look through the "Writing Annual Data to Oracle.R" script for more details
oracle.bio <- format.BIODATA.onesite("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2025/data/Vaughan 2025 biocharacteristics data.csv")

# You can get someone else to double check the data just to be safe
#saveRDS(oracle.bio, file = "biodata_for_checking.rds")

# Once you are happy with these data, upload to the GASPERA database.
# NOTE: make sure overwrite = FALSE
# dbWriteTable(
#   conn = channel,
#   schema = "GASPEREA",
#   value = oracle.bio,
#   name = "ALOSA_FISH_BIO_DATA",
#   date = TRUE,
#   row.names = FALSE,
#   overwrite = FALSE, ## NEVER CHANGE TO TRUE
#   append = TRUE
# )

```

## Cleaning / uploading count data to db
```{r}
# Go through the count data to double check things. Things to look for:
# - Check for really big or really small variances (maybe a keystroke error)
# - Gaps
# - Did we miss any counts, do we have the footage to redo it? 
# - Large number of downstream movement.

oracle.count <- format.COUNTDATA.onesite("C:/Users/graylo/Documents/Vaughan 2025 count data.csv")

# If you are happy with it, upload to the GASPERA database
# NOTE: make sure overwrite = FALSE
# dbWriteTable(
#   conn = channel,
#   schema = "GASPEREA",
#   value = oracle.count,
#   name = "ALOSA_VIDEO_COUNT_DATA",
#   date = TRUE,
#   row.names = FALSE,
#   ## NEVER CHANGE TO TRUE
#   overwrite = FALSE,
#   append = TRUE
# )
```

## Selecting scales
```{r}
# We use the ageing.selection.v1.1 script to select which scales we choose to
# process for ageing. It uses the daily.summary object from the
# onespecies.river.escapement script.
temp <- onespecies.river.escapement(
  "does_not_matter",
  fixtime = TRUE,
  downstream.migration = FALSE,
  database = TRUE,
  year = current_year,
  site = 2,
  channel = channel
)

# Get the biodata for the year from the db so it is all formated nicely
biodata_db <- dbReadTable(channel, "ALOSA_FISH_BIO_DATA")
biodata_db |> filter(SITE_ID == 2 & YEAR == current_year) -> biodata_tusket

# Since we do not collect data every day, we run the missing.days.R script
missing_days <- missing.days(biodata = biodata_tusket)

# This selects the scales to age
# Note: we don't put anything in the mergedays argument because this is only
# needed if weekly = FALSE
ageing.selection.test(
  countdata = temp,
  biodata = biodata,
  # This weights samples by the week and is used when weekends were not sampled
  # We use this at Tusket since we do not sample on weekends because we are
  # cool and based
  weekly = TRUE,
  year = current_year,
  missingdays = missing_days
  )

# This will output a csv file to the working directory
```

## VD species proportions
```{r species}
# Get the counts for the year for Lake Vaughan
counts <- onespecies.river.escapement(
  "lololololol",
  fixtime = TRUE,
  downstream.migration = FALSE,
  database = TRUE,
  year = current_year,
  site = 2,
  channel = channel
)

counts <- counts %>%
    mutate(across(everything(), ~ ifelse(is.nan(.), 0, .))) %>%
    mutate(across(everything(), ~ ifelse(is.na(.), 0, .)))

counts <- counts |> mutate(date = make_date(current_year, mon, day))

counts$year <- current_year

# This calculates the proportions of alewives and BBs in the escapement estimates
# based on the biodata where we took information on species proportions.
proportions <- split_species(
    year = current_year,
    siteID = 2,
    channel = channel
  )

counts_species <- counts |>
      mutate(date = make_date(year, mon, day)) |>
      ungroup() |>
      select(date, total, clow, chigh) |>
      inner_join(proportions, by = "date") |>
      mutate(B = total * BBprop, A = total - B) |>
      select(date, A, B) |>
      pivot_longer(cols = c(A, B),
                   names_to = "species",
                   values_to = "total") |>
      mutate(species = str_replace(species, "A", "Alewives")) |>
      mutate(species = str_replace(species, "B", "Bluebacks")) |>
      ungroup()

# This plot shows total counts of each species - this is not very useful
# because there are so few BBs
# counts_species |>
#   group_by(species) |>
#   ggplot(aes(date, total)) +
#   geom_path(aes(colour = species), size = 1) +
#   geom_point(aes(fill = species), shape = 21, colour = "black", size = 2)+
#   theme_bw() +
#   labs(
#     title = paste0(
#       "Escapement estimates for gaspereau by species for ",
#       current_year
#     ),
#     y = "Fish / day",
#     colour = "Species",
#     fill = "Species"
#   ) +
#   scale_x_date(date_labels = "%b %d", date_breaks = "3 days") +
#   scale_y_continuous(
#     limits = c(0, max(counts_species$total, na.rm = TRUE)),
#     breaks = seq(0, max(counts_species$total, na.rm = TRUE), by = 10000),
#     labels = scales::comma
#   ) +
#   theme(
#     axis.text.x = element_text(
#       angle = 45,
#       vjust = 1,
#       hjust = 1
#       ),
#     plot.background = element_rect(fill = "grey95"),
#     legend.background = element_blank(),
#     legend.key = element_blank(),
#     axis.title.x = element_blank(),
#     panel.grid.minor = element_blank()
#   )

# Here is a plot looking at percentages of species over time
# Get the proportions
counts_species |> 
  group_by(date) |> 
  mutate(percent = total / sum(total) * 100) |> 
  ungroup() |> 
  filter(!is.nan(percent)) -> species_props

species_props |> 
group_by(species) |>
  ggplot(aes(date, percent)) +
  geom_path(aes(colour = species), size = 1) +
  geom_point(aes(fill = species), shape = 21, colour = "black", size = 2)+
  theme_bw() +
  labs(
    title = paste0(
      "Proportions of escaped fish by species for ",
      current_year
    ),
    y = "Percentage (%)",
    colour = "Species",
    fill = "Species"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "3 days") +
  scale_y_continuous(
    limits = c(0, 100),
    breaks = seq(0, 100, 10)
  ) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      vjust = 1,
      hjust = 1
      ),
    plot.background = element_rect(fill = "grey95"),
    legend.background = element_blank(),
    legend.key = element_blank(),
    axis.title.x = element_blank(),
    panel.grid.minor = element_blank()
  )

```

## VD species split array
```{r}
# We can also use the twospecies.river.escapement function to get the species
# splits, but this has the added benefit of calculating CIs. We use the 
# proportions df we made above in the function
species_df <- twospecies.river.escapement(
  filename = "",
  year = current_year,
  site = 2,
  channel = channel,
  species.split = proportions
)

# Alewives
species_df[[1]]

# Bluebacks
species_df[[2]] -> BB

# Deal with the NaNs so that the plot is good
BB %>%
    mutate(across(everything(), ~ ifelse(is.nan(.), 0, .))) %>%
    mutate(across(everything(), ~ ifelse(is.na(.), 0, .))) -> BB

BB |>
  mutate(date = make_date(year = current_year, month = mon, day = day)) |> 
  ggplot(aes(date, total)) +
  geom_path(
    colour = "#7ECDBB",
    alpha = 0.9,
    linewidth = 1.25
    ) +
  geom_ribbon(
    aes(ymin = clow, ymax = chigh),
    fill = "#7ECDBB",
    colour = "grey90",
    alpha = 0.2
  ) +
  theme_bw() +
  labs(
    title = paste0("Gaspereau escapement at the Lake Vaughan fish ladder in ", current_year),
    x = "Date",
    y = "Fish / day"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 day") +
  scale_y_continuous(
    limits = c(0, max(BB$chigh)),
    breaks = seq(0, max(BB$chigh), by = 1e2),
    labels = scales::comma
  ) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      vjust = 1,
      hjust = 1
      ),
    axis.title.x = element_blank(),
    panel.grid.minor = element_blank(),
    #legend.background = element_blank(),
    plot.background = element_rect(fill = "grey95")
    )

# Cumulative plot of BBs
BB_cumulative <- BB |> mutate(cumulative_total = cumsum(total))

BB_cumulative |> 
  mutate(date = make_date(year = current_year, month = mon, day = day)) |> 
  ggplot(aes(date, cumulative_total)) +
  geom_path(colour = "purple", alpha = 0.8, linewidth = 1.75, na.rm = TRUE)+
  theme_bw() +
  labs(
    title = paste("Cumulative escapement of Blueback Herring at Lake Vaughan Dam in", current_year),
    x = "Date",
    y = "Total fish escaped",
    linetype = "Location"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 day") +
  scale_y_continuous(
    limits = c(0, max(BB_cumulative$cumulative_total)),
    breaks = seq(0, max(BB_cumulative$cumulative_total), 2e2),
    labels = scales::comma
  ) +
  theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.background = element_blank(),
    plot.background = element_rect(fill = "grey95"),
    legend.key = element_blank()
  ) +
  guides(colour = guide_legend(override.aes = list(linewidth = 2)))
```

# Species GLM

Only need to do this if you don't check for BBs every day during the late run.
In 2024 we did not check on weekends, so we needed to use this to model the BBs.
In 2025 we did check on weekends, so we do not need it.
```{r GLM_approach}
# We often have days that we could not sample biodata for various reasons. We
# can use padr::pad to identify which dates are missing and then fill with NAs.
pp <- padr::pad(proportions, interval = "day")

# To prevent trouble down the road, we create a column to store the "date" as 
# an integer value. When date is used in the model, it really doesn't have to be
# a date, it just has to be an integer. As dates can cause headaches with many
# functions etc. it is safe to use an integer instead.
pp$day_int <- 1:nrow(pp)

# We use a Poisson GLM as we are dealing with count data. This model is imperfect
# and we know this, but for our current requirements, it is fairly useful. We
# wouldn't want to use this model to extrapolate very far out. We train the model
# on the complete data set.
model <- glm(BB ~ day_int, data = pp, offset = log(total_fish), family = "poisson")

# We set the total_fish to 1 so that we get proportions when we predict the new
# data. We usually try to get 100 fish sampled per day in the biodata, but that
# doesn't always happen, especially at the tails of the season.
out <- data.frame(day_int = 1:nrow(pp), total_fish = rep(1,nrow(pp)))
out.pred <- predict(model, newdata = out, type = "link", se.fit = T)

# This is the inverse of the link function i.e. the inverse of the log i.e exp
ginv <- model$family$linkinv

# We use the inverse of the link function to calculate the expected proportion.
# Note, this is the same as doing exp(out.pred[[1]])
out.pred <- ginv(out.pred[[1]])

# Create a new column for the new predicted proportions of BBs
pp$BBpropnew<-NA

# Populate the dataframe with the predicted values
for(i in 1:nrow(pp)) {
  if (is.na(pp$BBprop[i]) == T) {
    pp$BBpropnew[i] <- out.pred[i]
    pp$BBprop[i] <- out.pred[i]
  }
}

# We can flag which data were observed and which were modeled
pp$observation <- is.na(pp$total_fish)
pp <- pp |> mutate(observation = if_else(observation == TRUE, "Modeled", "Observed"))

# Check the plot
pp |> 
  ggplot(aes(date, BBprop))+
  geom_path(colour = "grey50", linewidth = 0.5, linetype = "dashed")+
  geom_point(aes(fill = observation), shape = 21, colour = "black", size = 2)+
  theme_bw()+
  labs(
    title = "Proportion of blueback herring in biodata with values for missing dates from GLM",
    y = "Proportion of fish",
    fill = "Estimate"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = .1),
    labels = scales::comma
  ) +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    plot.background = element_rect(fill = "grey95"),
    axis.title.x = element_blank(),
    legend.background = element_blank(),
    legend.key = element_blank()
  )

```

## Species escapement VD
```{r blueback_escapement_vd}
# Combine the Lake Vaughan escapement estimates with the proportions dataframe
# where we calculated the proportions of bluebacks for each date
result <- left_join(pp, counts, by = "date")
result <- result |> 
  select(date, BBprop, total) |> 
  mutate(
    BB = BBprop * total,
    A = total - BB
    ) |>
  select(-BBprop, -total) |> 
  pivot_longer(
    cols = c(A, BB),
    names_to = "Species",
    values_to = "Total"
    )

result |> 
  ggplot(aes(date, Total))+
  geom_path(aes(colour = Species), size = 1)+
  geom_point(aes(fill = Species), shape = 21, colour = "black", size = 2)+
  theme_bw()+
  labs(
    title = "Estimated escapement for alewives and bluebacks at Lake Vaughan for 2024",
    y = "Fish / day"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  scale_y_continuous(
    limits = c(0, max(result$Total)),
    breaks = seq(0, max(result$Total), by = 10000),
    labels = scales::comma
  ) +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    plot.background = element_rect(fill = "grey95"),
    axis.title.x = element_blank(),
    legend.background = element_blank(),
    legend.key = element_blank()
  )

```

## Species counts VD + PH
```{r total_counts}
# I use the CSVs for counts here because I can only get 2024 data to run from
# the database without throwing an error, 2022 and 2023 do not work
path_to_counts_CSV <- "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2024/Data Sheets/Vaughan 2024 count data.csv"

vd_split <- split_species(
  year = current_year,
  siteID = 2,
  channel = channel,
  biodata_CSV = NULL
  )

counts_vd <- twospecies.river.escapement(
  path_to_counts_CSV,
  fixtime = TRUE,
  downstream.migration = FALSE,
  database = FALSE,
  year = current_year,
  site = 2,
  channel = channel,
  species.split = vd_split
)

# Got to grab each of the lists within counts_vd which get created by 
# twospecies.river.escapement and then add a column to each indicating which
# species they represent (A = Alewives, B = Bluebacks)
vd_alewives <- as_tibble(counts_vd[[1]])
vd_alewives$species = "A"

vd_bluebacks <- as_tibble(counts_vd[[2]])
vd_bluebacks$species = "B"

counts_vd <- rbind(vd_alewives, vd_bluebacks)
counts_vd$location <- "Lake Vaughan"

# We use the split for Vaughan Dam here because there are no counts made at PH
path_to_counts_CSV <- "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2024/Data Sheets/Powerhouse 2024 count data.csv"

counts_ph <- twospecies.river.escapement(
  path_to_counts_CSV,
  fixtime = TRUE,
  downstream.migration = FALSE,
  database = FALSE,
  year = current_year,
  site = 14,
  channel = channel,
  species.split = vd_split
)

ph_alewives <- as_tibble(counts_ph[[1]])
ph_alewives$species = "A"

ph_bluebacks <- as_tibble(counts_ph[[2]])
ph_bluebacks$species = "B"

counts_ph <- rbind(ph_alewives, ph_bluebacks)
counts_ph$location <- "Powerhouse"

# Join these counts into a single dataframe so we can add them
counts_joined <- rbind(counts_vd, counts_ph)

counts <- counts_joined %>%
  mutate(
    across(
      everything(), ~ replace_na(., 0)
      )
    )

# Convert species counts to a format that we can easily copy into the 
# "daily_counts_all_years_all_rivers.csv" file for the plots in the 
# Tusket River Annual Summary Document.
counts_for_csv <- counts |>
  group_by(species) |>
  summarise(total_escapement = sum(total))

counts_for_csv <- counts_for_csv |> 
  bind_rows(
    counts_for_csv |>
      summarise(total_escapement = sum(total_escapement)) |> 
      mutate(species = "All")
  )

```

## Species catch

These numbers are required for the `Tusket River Annual Summary` document that is
made every year. They are estimates of the number of bluebacks caught in the Tusket.
The numbers here are used to fill in data in `river summaries all years.csv` that
from which `Tusket River Annual Summary Document_Script.Rmd` pulls its data when writing
the document.

```{r catch_data}
inspection <- readxl::read_xlsx(
  "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2022/Data Sheets/2022 Yarmouth County Catch Inspection.xlsx",
  sheet = "Sheet1"
  )

inspection <- inspection |> mutate(date = make_date(year = year, month = mon, day = day))
may_13_2022_prop <- inspection |> 
  filter(date >= "2022-05-13" & date <= "2022-05-15" & method == "gill" & river == "Tusket") |> 
  summarize(BBprop = sum(n.blueback) / sum(n.alewife))

may_20_2022_prop <- inspection |> 
  filter(date >= "2022-05-20" & date <= "2022-05-22" & method == "gill" & river == "Tusket") |> 
  summarize(BBprop = sum(n.blueback) / sum(n.alewife))

```

```{r blueback_catch}
# I THINK I NEED TO REDO THIS WHOLE SECTION AND USE OFFSETS OF 7 AND 3 DAYS LIKE 
# I HAVE DONE FOR OTHER ANALYSES
# 
# HOW MANY WEEKS SHOULD WE CONSIDER THE DIPPERS AS HAVING ACCESS TO BB?
# 
# 
# 
# This section will need to be edited for each year. The way this is currently
# set up requires you to identify which of the last weeks of May you want to
# calculate blueback catch for. This will be based on when bluebacks were
# detected. In the example below, we only used gill net data as there were no
# bluebacks found at the dipping stands when George and Mark sampled their
# catches in 2022.
load("C:/Users/graylo/Documents/data/marfis_pull.Rdata")
source("R:/Science/Population Ecology Division/DFD/Alosa/MARFISSCI/MARFIS.error.cleaner.R")
catch <- convert.KGS(catch)

# Total tusket catch
tusket_total_catch <- catch |>
  filter(YEAR == 2024 & RIVERNAME_CLEANED == "TUSKET") |> 
  # We divide by the conversion factor to convert mass to count
  summarise(total_catch = sum(KGS) / .240)

# I checked the logbooks for the Tusket licences from the filtered data above, and they all look good for 2024. There were a few small mistakes that I won't correct here.
tusket_gill <- catch |>
  filter(YEAR == 2024 & RIVERNAME_CLEANED == "TUSKET" & GEAR_DESCRIPTION == "GILL NET (SET OR FIXED)")

# Now for the dippers
tusket_gill <- catch |>
  filter(YEAR == 2024 & RIVERNAME_CLEANED == "TUSKET" & GEAR_DESCRIPTION == "GILL NET (SET OR FIXED)")

tusket_gill <- as_tibble(tusket_gill)

# Get landings per week for last two weeks for gill netters only
week_1 <- tusket_gill |> filter(FV_DATE_FISHED >= "2024-05-17" & FV_DATE_FISHED <= "2024-05-20") |> summarise(total_kg = sum(KGS))
week_2 <- tusket_gill |> filter(FV_DATE_FISHED >= "2024-05-24") |> summarise(total_kg = sum(KGS))

# Convert landings to individual fish using 0.240 kilograms per fish 
# (i.e. a fish has a mass of 240 g) and make tibble of all estimates
bb_catch <- tibble(
  week1 = round(week_1$total_kg * may_13_2022_prop$BBprop / .240),
  week2 = round(week_2$total_kg * may_20_2022_prop$BBprop / .240)
  )

bb_catch <- bb_catch |> mutate(total_bb_catch = rowSums(across(everything())))

bb_catch <- bb_catch$total_bb_catch
```

```{r blueback_exploitation}
pp$date <- ymd(pp$date)
counts <- counts |> mutate(date = make_date(year = year, month = mon, day = day))
result <- left_join(pp, counts, by = "date")

result <- result |> 
  select(date, BBprop, total) |> 
  mutate(
    BB = BBprop * total,
    A = total - BB
  ) |>
  select(-BBprop, -total) |> 
  pivot_longer(
    cols = c(A, BB),
    names_to = "Species",
    values_to = "Total"
  )

bb_total <- result |> 
  group_by(Species) |> 
  summarise(sum = sum(Total))

# Exploitation rate
bb_exploitation <- bb_catch / (bb_total$sum[2] + bb_catch)
```

```{r summary_table}
# Label counts from Lake Vaughan
counts$location <- "Lake Vaughan"
counts <- counts |> select(total, location)
total_count_vd <- sum(counts$total)

# PH counts
ph_counts <- onespecies.river.escapement(
  fixtime = FALSE,
  downstream.migration = FALSE,
  database = TRUE,
  year = current_year,
  site = 14,
  channel = channel
)
ph_counts$location <- "Powerhouse"
ph_counts <- ph_counts |> select(total, location)
total_count_ph <- sum(ph_counts$total)

# Combine counts from both sites
counts <- rbind(counts, ph_counts)
total_count <- sum(counts$total)

df <- tibble(
  tusket_escapement = total_count,
  VD_escapement = total_count_vd,
  PH_escapement = total_count_ph,
  VD_bb_escapement = bb_total$sum[2],
  tusket_catch = round(tusket_total_catch$total_catch),
  tusket_catch_bb = bb_catch
)

df <- df |> pivot_longer(cols = everything(), names_to = "variable", values_to = "count")
df <- df |> mutate(count = round(count))
df <- df |> mutate(count = scales::comma(count))
```

```{r summary_table2, results='asis'}
library(gt)

gt_table <-
  df |> 
  gt() |> 
  tab_header(title = "Season summary") |>
  cols_label(
    variable = "Observation",
    count = "Count"
  )

gt_table
```

