The purpose of this document is to haul relevant data for use at the advisory
committee meetings for the Yarmouth / Shelburne region for 2025.

# SETUP
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,     # Suppress warnings
  message = FALSE     # Suppress messages
)
```

```{r libraries_et_al}
library(tidyverse)
library(DBI)
library(ROracle)

# Set up the connection to the database
channel <- dbConnect(
  DBI::dbDriver("Oracle"),
  username = "GASPEREA",
  password = "_REMOVED",
  dbname = "PTRAN",
  believeNRows = FALSE
)

# Source all the functions so we can use them
source("~/git/ALOSA.functions/functions/sourcery.R")
sourcery()
```

# LOAD DATA
```{r load_MARFIS_data, include=FALSE, warning=FALSE}
# Do a new MARFIS pull
#source("~/git/ALOSA.functions/MARFIS_all in one.R")

# save(
#   catch,
#   didnotfish,
#   licencerenewals,
#   file = "C:/Users/graylo/Documents/data/marfis_pull.Rdata"
#   )

# Load in the data
load("C:/Users/graylo/Documents/data/marfis_pull.Rdata")

# Clean the data
source("~/MARFIS.error.cleaner.R")

# Convert weights to kilograms
catch <- convert.KGS(catch)
catch <- subset(catch, select = -FV_WEIGHT)
catch <- catch |> rename(FV_WEIGHT = KGS)

# Get the licence data provided by RM
licences <- readxl::read_excel("R:/Science/Population Ecology Division/DFD/Alosa/MARFISSCI/Gaspereau Licences Area Project.xlsx")
```

# REPORTING RATE

NOTE: cahnge this so that it just uses the MARFIS pull data and stop relying on
the CSV from RM as it seems to be out-dated i.e. 
Get the licences registered from our last point of data (2023) for YS
```{r reporting_rate}
# Total licences renewed in 2023 for NB
licences |> 
  select(STATUS, LICENCE_ID = `Licence Id`, RENEWED = Renewed, COUNTY = `County Name 1`, PROVINCE = `Province Abbrev (English)`, LICENCE_TYPE = `Licence Type Description (English)`) |> 
  filter(RENEWED == 2023 & PROVINCE == "NS") |> 
  filter(COUNTY == "YARMOUTH" | COUNTY == "SHELBURNE") |> 
  distinct(LICENCE_ID, .keep_all = TRUE) |> 
  arrange(LICENCE_ID) |> 
  mutate(LICENCE_ID = as.factor(LICENCE_ID)) -> ys_licences_2023

ys_licences_2023
# Based on some sleuthing, all the licences appear to be good, so no need to
# filter out any inactive licences etc
```

Break down licence registrations by county
```{r}
ys_licences_2023 |> 
  count(COUNTY)
```

See which of these submitted catch records in 2024 (i.e. subset catch)
```{r}
# This is a list of the licences that reported catch in 2024
catch |> 
  filter(YEAR == 2024) |> 
  select(LICENCE_ID, YEAR, RIVERNAME = RIVERNAME_CLEANED, COUNTY, PROVINCE, FV_WEIGHT) |> 
  filter(LICENCE_ID %in% ys_licences_2023$LICENCE_ID) |> 
  distinct(LICENCE_ID) -> ys_licences_catch_2024

ys_licences_catch_2024$status <- "Catch"

ys_licences_catch_2024
```

Where were the licences that reported catch fishing?
```{r}
catch |> 
  filter(YEAR == 2024) |> 
  filter(LICENCE_ID %in% ys_licences_catch_2024$LICENCE_ID) |> 
  distinct(LICENCE_ID, .keep_all = TRUE) |> 
  rename(RIVERNAME = RIVERNAME_CLEANED) |> 
  # count(RIVERNAME_CLEANED) |> 
  # arrange(desc(n))
  mutate(
  AREA = case_when(
    RIVERNAME == "TUSKET" ~ "TUSKET",
    RIVERNAME == "ANNIS" ~ "ANNIS",
    RIVERNAME == "KIACK BROOK" | RIVERNAME == "EEL LAKE" ~ "KIACK BROOK / EEL LAKE",
    RIVERNAME == "ARGYLE" ~ "ARGYLE",
    RIVERNAME == "BARRINGTON" | RIVERNAME == "ROSEWAY" | RIVERNAME == "PORT SAXON" ~ "SHELBURNE COUNTY",
    RIVERNAME == "MILTON" 
    | RIVERNAME == "YARMOUTH HARBOUR" 
    | RIVERNAME == "DUNN" 
    | RIVERNAME == "ARCADIA" 
    | RIVERNAME == "PORT MAITLAND" 
    | RIVERNAME == "SHORT BEACH" 
    | RIVERNAME == "ISLAND POND" 
    | RIVERNAME == "GREENVILLE" 
    | RIVERNAME == "PEMBROKE" 
    | RIVERNAME == "MELBOURNE" 
    | RIVERNAME == "CHEGGOGIN" 
    | RIVERNAME == "STINK PLANT" 
    | RIVERNAME == "BEAVER RIVER" ~ "YARMOUTH COUNTY (OTHER)",
    TRUE ~ NA_character_
    )
  ) |> 
  group_by(AREA) |> 
  count() |> 
  arrange(desc(n))
```

See who reported catch by county
```{r}
catch |> 
  filter(YEAR == 2024) |> 
  filter(LICENCE_ID %in% ys_licences_catch_2024$LICENCE_ID) |> 
  distinct(LICENCE_ID, .keep_all = TRUE) |> 
  count(COUNTY) |> 
  arrange(desc(n))
```

See which of these submitted DNF (i.e. subset didnotfish)
```{r}
didnotfish |> 
  filter(YEAR == 2024 & NIL_REPORT_FLAG == "Y") |> 
  arrange(LICENCE_ID) |> 
  select(LICENCE_ID) -> dnf

# Get a summary of how many licences reported catch, dnf, or did not report
ys_licences_2023 |> filter(LICENCE_ID %in% dnf$LICENCE_ID) -> ys_licences_dnf

ys_licences_dnf$status <- "DNF"

ys_licences_dnf
```

See who DNF by county
```{r}
ys_licences_dnf |> 
  count(COUNTY)
```

Combine the licences that reported catch or DNF into one df
```{r}
ys_licences_catch_2024 |> 
  full_join(ys_licences_dnf) -> ys_licences_reported

ys_licences_reported
```


Assume the remainder of the licences Did Not Report (DNR)
```{r}
ys_licences_2023 |> 
  anti_join(ys_licences_reported, by = "LICENCE_ID") |> 
  distinct() |> 
  mutate(status = "DNR") -> ys_licences_DNR

ys_licences_DNR
```
See DNR by county
```{r}
ys_licences_DNR |> count(COUNTY)
```

Break down total proportions for each reporting status
```{r}
ys_licences_reported |> 
  full_join(ys_licences_DNR) -> ys_lic_df

ys_lic_df |> 
  count(status) |> 
  mutate(proportion = n / sum(n)) |> 
  arrange(desc(proportion))
```

# ESCAPEMENT

```{r escapement plot}
# These data show historic escapement estimates at the Lake Vaughan Dam. These
# values are taken from the values hard coded into the Annual Tusket Assessment
# Rmd script in the ALOSA functions folder
esc <- read_csv(
  "C:/Users/graylo/Documents/GitHub/ac-tusket-2025/historical_escapement.csv",
  col_types = cols(
    year = col_character(),
    total = col_integer()
    )
  )

esc |> 
  arrange(year) |> 
  mutate(total = total / 1e6) |> 
  ggplot(aes(year, total)) +
  geom_bar(colour = "#2F4F4F", fill = "#66CDAA", stat = "identity") +
  labs(
    title = "Estimated spawning escapement for alewives on Tusket River",
    y = "Escapement (millions of fish)"
    ) +
  scale_y_continuous(
    limits = c(0, 3.5),
    breaks = seq(0, 3.5, 0.5)
  ) +
  geom_hline(yintercept = 2.086, linetype = "dashed") + #cautious
  geom_hline(yintercept = 3.099, linetype = "dashed") + #healthy
  annotate("text", x = 5.5, y = 1.75, label = "Critical", size = 5) +
  annotate("text", x = 5.5, y = 2.6,  label = "Cautious", size = 5) +
  annotate("text", x = 5.5, y = 3.4,  label = "Healthy", size = 5) +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    plot.title = element_text(size = 16),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 14),
    plot.background = element_blank()
  ) +
  xlab("")

ggsave(
  "escapement.png",
  plot = last_plot(),
  width = 9,
  height = 6,
  units = c("in"),
  bg = "transparent"
  )
```
```{r two_species_escapement}
# Split species - impute data for missing days
proportions <- split_species(year = 2024, siteID = 2, channel = channel)
 
# Use padr to "pad" the data set with NAs where we are missing observations. 
pp <- padr::pad(proportions, interval = "day")
pp <- pp |> select(date, BBprop)
pp <- imputeTS::na_interpolation(pp |> select(date, BBprop)) # Linear interpolation
pp$mon<-month(pp$date)
pp$day<-day(pp$date)
 
# Get escapement data for VD and PH
vd_2024 <- twospecies.river.escapement(
  "dummy",
  fixtime = TRUE,
  downstream.migration = FALSE,
  database = TRUE,
  year = 2024,
  site = 2,
  species.split=pp,
  channel = channel
)

vd_2024_ale <- as.data.frame(vd_2024[1])
vd_2024_bb <- as.data.frame(vd_2024[2])

ph_2024 <- twospecies.river.escapement(
  "dummy",
  fixtime = FALSE,
  downstream.migration = FALSE,
  database = TRUE,
  year = 2024,
  site = 14,
  species.split = pp,
  channel = channel
)

ph_2024_ale <- as.data.frame(ph_2024[1])
ph_2024_bb <- as.data.frame(ph_2024[2])

# Make a data frame containing the totals
totals <- data.frame(
  site = c("Lake Vaughan", "Powerhouse"),
  total_a = c(sum(vd_2024_ale$total), sum(ph_2024_ale$total)),
  total_bb = c(sum(vd_2024_bb$total), sum(ph_2024_bb$total))
)

# Add info and combine data
vd_2024_ale$species <- "Alewife"
vd_2024_ale$location <- "Lake Vaughan"
ph_2024_ale$species <- "Alewife"
ph_2024_ale$location <- "Powerhouse"

vd_2024_bb$species <- "Blueback"
vd_2024_bb$location <- "Lake Vaughan"
ph_2024_bb$species <- "Blueback"
ph_2024_bb$location <- "Powerhouse"

all <- rbind(vd_2024_ale, vd_2024_bb, ph_2024_ale, ph_2024_bb)

# Remove NAs and NaNs in the clow and chigh
all <- all %>% 
  mutate(across(everything(), ~ ifelse(is.nan(.), 0, .))) %>%
  mutate(across(everything(), ~ ifelse(is.na(.), 0, .)))

# Add dates so we can plot the thing
all <- all |> 
  mutate(date = make_date(year = 2024, month = mon, day = day)) |> 
  select(date, location, species, total, clow, chigh)

# Plot it
all |>
  group_by(location) |>
  ggplot(aes(date, total, colour = species, fill = species))+
  facet_wrap(~location, scales = "fixed", ncol = 1)+
  geom_path(
    data = all |> filter(location == "Lake Vaughan"),
    #aes(colour = location),
    alpha = 0.7,
    linewidth = 1.25
  )+
  geom_ribbon(
    data = all |> filter(location == "Lake Vaughan"),
    aes(ymin = clow, ymax = chigh),
    alpha = 0.5
  )+
  geom_path(
    data = all |> filter(location == "Powerhouse"),
    #aes(colour = location),
    alpha = 0.7,
    linewidth = 1.25
  )+
  geom_ribbon(
    data = all |> filter(location == "Powerhouse"),
    aes(ymin = clow, ymax = chigh),
    alpha = 0.5
  )+
  theme_bw() +
  labs(
    title = "Daily escapement estimates for gaspereau on the Tusket River in 2024",
    y = "Escapement (fish / day)",
    colour = "Species",
    fill = "Species"
    )+
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week")+
  scale_y_continuous(
    limits = c(0, max(all$chigh)),
    breaks = seq(0, max(all$chigh), by = 2e4),
    labels = scales::comma
  )+
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )
  #+ coord_cartesian(xlim = as.Date(c("2024-05-20", "2024-06-17")), ylim = c(0, 1e4))

ggsave("daily_tusket_zoomed_2024.png", plot = last_plot(), width = 10, height = 6, units = "in", dpi = 300, bg = "transparent")


```

# STOCK STATUS

```{r stock_status}
# Here is the status plot
# 
# Get the landings data and convert them
# load("C:/Users/graylo/Documents/data/marfis_pull.Rdata")
# source("C:/Users/graylo/Documents/MARFIS.error.cleaner.R")
# catch <- convert.KGS(catch)
# catch <- subset(catch, select = -FV_WEIGHT)
# catch <- catch |> rename(FV_WEIGHT = KGS)

# Get landings for tusket by day, figure out the singletons, and then get them into
# a dataframe for years >= 2021
tusket <- catch |> 
  filter(YEAR >= 2021) |>
  filter(RIVERNAME_CLEANED == "TUSKET") |> 
  select(LICENCE_ID, RIVERNAME_CLEANED, YEAR, MONTH, DAY, FV_DATE_FISHED, FV_GEAR_CODE, GEAR_DESCRIPTION, FV_HOURS_FISHED, FV_WEIGHT) |> 
  mutate(
    GEAR_DESCRIPTION = case_when(
      str_detect(GEAR_DESCRIPTION, "DIP") ~ "DIP NET",
      TRUE ~ GEAR_DESCRIPTION
      ),
    GEAR_DESCRIPTION = case_when(
      str_detect(GEAR_DESCRIPTION, "SQUARE") ~ "GILL NET (SET OR FIXED)",
      TRUE ~ GEAR_DESCRIPTION
      )
    ) |>
  mutate(
    FV_DATE_FISHED = make_date(year = YEAR, month = MONTH, day = DAY),
    YEAR = as.factor(YEAR),
    GEAR_DESCRIPTION = as.factor(GEAR_DESCRIPTION)
    ) |> 
  group_by(YEAR) |> 
  summarise(TOTAL_CAUGHT = sum(FV_WEIGHT) / .240)

# Get the escapement
get_escapement <- function(year){
  df <- onespecies.river.escapement(
  "dummy",
  fixtime = FALSE,
  database = TRUE,
  year = year,
  site = 2,
  channel = channel
  )
  
  df$year <- year
  df$date <- make_date(year = df$year, month = df$mon, day = df$day)
  return(df)
}

escapement_df <- data.frame()
for (year in c(2021, 2022, 2023, 2024)){
  df <- get_escapement(year)
  escapement_df <- rbind(escapement_df, df)
}
escapement_df <- escapement_df |> 
  select(year, date, total_esc = total, sd, clow, chigh)

escapement_df$location <- "Lake Vaughan"

escapement_df

# The powerhouse data worked this way for me, so I'll do it from CSV for now
ph_2022 <- onespecies.partial.river.escapement(
  "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2022/Data Sheets/Counts/Powerhouse Count Sheet Cleaned 2022.csv",
  fixtime = FALSE,
  database = FALSE,
  2022,
  14,
  channel
)

ph_2023 <- onespecies.river.escapement(
  "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2023/Data Sheets/Powerhouse 2023 count data1.csv",
  fixtime = FALSE,
  downstream.migration = FALSE,
  database = FALSE,
  2023,
  14,
  channel
)

ph_2024 <- onespecies.river.escapement(
  "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2024/Data Sheets/Powerhouse 2024 count data.csv",
  fixtime = TRUE,
  downstream.migration = FALSE,
  database = FALSE,
  2024,
  14,
  channel
)

ph_2022$year <- "2022"
ph_2023$year <- "2023"
ph_2024$year <- "2024"

ph <- rbind(ph_2022, ph_2023, ph_2024)
ph$location <- "Powerhouse"
ph$date <- make_date(year = ph$year, month = ph$mon, day = ph$day)
ph <- ph |> select(year, date, total_esc = total, sd, clow, chigh, location)

# Bind the sites data together
escaped <- rbind(escapement_df, ph)

# Group by date and sum the escapement
escaped <- escaped |> 
  group_by(year) |> 
  summarise(total_esc = sum(total_esc)) |> 
  mutate(YEAR = as.factor(year)) |> 
  rename(TOTAL_ESC = total_esc) |> 
  select(YEAR, TOTAL_ESC)

# Join the catch data and escapement data into a single table
status <- tusket |>
  left_join(escaped, by = "YEAR") |> 
  group_by(YEAR) |>
  mutate(ER = TOTAL_CAUGHT / (TOTAL_CAUGHT + TOTAL_ESC))

# Make the status plot
status <- status |> 
  mutate(ESC_USR = TOTAL_ESC / 3.098e6) |> 
  mutate(MU = ER / 0.53)

status |> 
  ggplot(aes(ESC_USR, MU, shape = YEAR)) +
  geom_point(size = 4, colour = "grey30")+
  theme_bw() +
  labs(
    title = "Stock status",
    y = expression("µ / µ"["RR"]),
    x = expression("Esc / Esc"["USR"]),
    shape = "YEAR"
    ) +
  scale_x_continuous(limits = c(0, 1.6), breaks = seq(0, 1.6, 0.2)) +
  scale_y_continuous(limits = c(0, 1.6), breaks = seq(0, 1.6, 0.2)) +
  geom_vline(xintercept = 2.086e6 / 3.098e6, linetype = "dotted", size = 3/5) +
  geom_vline(xintercept = 1, linetype = "dotted", size = 3/5) +
  geom_hline(yintercept = 0.35 / 0.53, linetype = "dotted", size = 3/5) +
  geom_hline(yintercept = 1, linetype = "dotted", size = 3/5) +
  annotate("text", x = 0.32, y = 1.6,  label = "Critical", size = 5) +
  annotate("text", x = 0.84, y = 1.6,  label = "Cautious", size = 5) +
  annotate("text", x = 1.4,  y = 1.6,  label = "Healthy", size = 5) +
  annotate("text", x = 1.6,  y = 1.3,  label = "Over \n exploited", angle = 90, size = 5) +
  annotate("text", x = 1.6,  y = 0.82, label = "Fully \n exploited", angle = 90, size = 5) +
  annotate("text", x = 1.6,  y = 0.3,  label = "Partially \n exploited", angle = 90, size = 5) +
  theme(
    plot.title = element_text(size = 18),
    axis.title.y = element_text(size = 16),
    axis.title.x = element_text(size = 16),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 12)
  )

ggsave("status.png", plot = last_plot(), width = 9, height = 6, units = c("in"))
```
# LANDINGS

```{r}
# This is the break down of all the licences reporting catch and their location,
# but grouped for each river
catch |> 
  select(LICENCE_ID, YEAR, RIVERNAME = RIVERNAME_CLEANED, COUNTY, PROVINCE) |> 
  filter(YEAR == 2024 & (COUNTY == "YARMOUTH COUNTY" | COUNTY == "SHELBURNE COUNTY")) |> 
  distinct(LICENCE_ID, .keep_all = TRUE) |> 
  group_by(RIVERNAME, COUNTY) |> 
  count() |> 
  arrange(desc(n))
```

```{r}
# This is the break down of all the licences reporting catch and their location,
# but grouped for each arbitrary area
catch |> 
  select(LICENCE_ID, YEAR, RIVERNAME = RIVERNAME_CLEANED, COUNTY, PROVINCE) |> 
  filter(YEAR == 2024 & (COUNTY == "YARMOUTH COUNTY" | COUNTY == "SHELBURNE COUNTY")) |> 
  distinct(LICENCE_ID, .keep_all = TRUE) |> 
  mutate(
    AREA = case_when(
      RIVERNAME == "TUSKET" ~ "TUSKET",
      RIVERNAME == "ANNIS" ~ "ANNIS",
      RIVERNAME == "KIACK BROOK" | RIVERNAME == "EEL LAKE" ~ "KIACK BROOK / EEL LAKE",
      RIVERNAME == "ARGYLE" ~ "ARGYLE",
      RIVERNAME == "BARRINGTON" | RIVERNAME == "ROSEWAY" | RIVERNAME == "PORT SAXON" ~ "SHELBURNE COUNTY",
      RIVERNAME == "MILTON" | RIVERNAME == "YARMOUTH HARBOUR" | RIVERNAME == "DUNN" | RIVERNAME == "ARCADIA" | RIVERNAME == "PORT MAITLAND" | RIVERNAME == "SHORT BEACH" | RIVERNAME == "ISLAND POND" | RIVERNAME == "GREENVILLE" | RIVERNAME == "PEMBROKE" | RIVERNAME == "MELBOURNE" | RIVERNAME == "CHEGGOGIN" | RIVERNAME == "STINK PLANT" ~ "YARMOUTH COUNTY (OTHER)",
      TRUE ~ NA_character_
    )
  ) |> 
  group_by(AREA) |> 
  count() |> 
  arrange(desc(n))
```

```{r}
# This groups the different rivers in Yar / Shel into different arbitrary 
# regions that are useful when we present at the AC meeting
catch |> 
  select(LICENCE_ID, YEAR, RIVERNAME = RIVERNAME_CLEANED, COUNTY, FV_WEIGHT) |> 
  filter(YEAR == 2024 & (COUNTY == "YARMOUTH COUNTY" | COUNTY == "SHELBURNE COUNTY")) |> 
  group_by(RIVERNAME) |> 
  summarise(FV_WEIGHT = round(sum(FV_WEIGHT))) |> 
  mutate(COUNT = round(FV_WEIGHT / 0.240)) |> 
  mutate(
    AREA = case_when(
      RIVERNAME == "TUSKET" ~ "TUSKET",
      RIVERNAME == "ANNIS" ~ "ANNIS",
      RIVERNAME == "KIACK BROOK" | RIVERNAME == "EEL LAKE" ~ "KIACK BROOK / EEL LAKE",
      RIVERNAME == "ARGYLE" ~ "ARGYLE",
      RIVERNAME == "BARRINGTON" | RIVERNAME == "ROSEWAY" | RIVERNAME == "PORT SAXON" ~ "SHELBURNE COUNTY",
      RIVERNAME == "MILTON" | RIVERNAME == "YARMOUTH HARBOUR" | RIVERNAME == "DUNN" | RIVERNAME == "ARCADIA" | RIVERNAME == "PORT MAITLAND" | RIVERNAME == "SHORT BEACH" | RIVERNAME == "ISLAND POND" | RIVERNAME == "GREENVILLE" | RIVERNAME == "PEMBROKE" | RIVERNAME == "MELBOURNE" | RIVERNAME == "CHEGGOGIN" | RIVERNAME == "STINK PLANT" ~ "YARMOUTH COUNTY (OTHER)",
      TRUE ~ NA_character_
    )
  ) |> 
  group_by(AREA) |> 
  summarise(FV_WEIGHT = sum(FV_WEIGHT), COUNT = sum(COUNT)) |> 
  arrange(desc(COUNT)) |> 
  mutate(WEIGHT_MT = round(FV_WEIGHT / 1000), WEIGHT_KG = round(FV_WEIGHT, -3), COUNT = round(COUNT, -3)) |> 
  select(-FV_WEIGHT)
```

```{r landings_area}
# This is the break down of all the YS landings for each year since 2019

catch |> 
  select(LICENCE_ID, YEAR, RIVERNAME = RIVERNAME_CLEANED, COUNTY, FV_WEIGHT) |> 
  filter(YEAR >= 2019 & (COUNTY == "YARMOUTH COUNTY" | COUNTY == "SHELBURNE COUNTY")) |> 
  group_by(RIVERNAME, YEAR) |> 
  summarise(FV_WEIGHT = round(sum(FV_WEIGHT))) |> 
  mutate(COUNT = round(FV_WEIGHT / 0.240)) |> 
  mutate(
    AREA = case_when(
      RIVERNAME == "TUSKET" ~ "TUSKET",
      RIVERNAME == "ANNIS" ~ "ANNIS",
      RIVERNAME == "KIACK BROOK" | RIVERNAME == "EEL LAKE" ~ "KIACK BROOK / EEL LAKE",
      RIVERNAME == "ARGYLE" ~ "ARGYLE",
      RIVERNAME == "BARRINGTON" | RIVERNAME == "ROSEWAY" | RIVERNAME == "PORT SAXON" ~ "SHELBURNE COUNTY",
      RIVERNAME == "MILTON" | RIVERNAME == "YARMOUTH HARBOUR" | RIVERNAME == "DUNN" | RIVERNAME == "ARCADIA" | RIVERNAME == "PORT MAITLAND" | RIVERNAME == "SHORT BEACH" | RIVERNAME == "ISLAND POND" | RIVERNAME == "GREENVILLE" | RIVERNAME == "PEMBROKE" | RIVERNAME == "MELBOURNE" | RIVERNAME == "CHEGGOGIN" | RIVERNAME == "STINK PLANT" ~ "YARMOUTH COUNTY (OTHER)",
      TRUE ~ NA_character_
    )
  ) |> 
  group_by(AREA, YEAR) |> 
  summarise(FV_WEIGHT = sum(FV_WEIGHT), COUNT = sum(COUNT)) |> 
  arrange(desc(COUNT)) |> 
  mutate(WEIGHT_MT = round(FV_WEIGHT / 1000), WEIGHT_KG = round(FV_WEIGHT, -3), COUNT = round(COUNT, -3)) |> 
  select(-FV_WEIGHT) |> 
  mutate(YEAR = as.factor(YEAR)) |> 
  drop_na(AREA) |> 
  ungroup() |> # Ungroup here so that the refactoring of the RIVERNAME works
  mutate(AREA = fct_reorder(AREA, WEIGHT_MT, .fun = max, .desc = TRUE)) -> ys_catch

ys_catch |> 
  ggplot(aes(YEAR, WEIGHT_MT, fill = AREA)) +
  geom_col(position = position_dodge(), colour = "#2F4F4F") +
  labs(
    title = "Annual reported gaspereau landings for Yarmouth / Shelburne",
    y = "Landings (mt)"
    ) +
  # scale_y_continuous(
  #   limits = c(0, 3e2),
  #   breaks = seq(0, 3e2, 50),
  #   labels = scales::label_comma()
  # ) +
  theme_bw() +
  theme(
    legend.position = "right",
    plot.title = element_text(size = 16),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    panel.grid.minor = element_blank()
    ) +
  xlab("")

ggsave("landings_ys_areas.png", plot = last_plot(), width = 9, height = 6, units = "in", dpi = 300)

ys_catch |> filter(YEAR == 2024)
```

```{r landings_county}
# This is the break down of all the YS landings for each year since 2019

catch |> 
  select(LICENCE_ID, YEAR, RIVERNAME = RIVERNAME_CLEANED, COUNTY, FV_WEIGHT) |> 
  filter(YEAR >= 2019 & (COUNTY == "YARMOUTH COUNTY" | COUNTY == "SHELBURNE COUNTY")) |> 
  group_by(COUNTY, YEAR) |> 
  summarise(FV_WEIGHT = round(sum(FV_WEIGHT))) |> 
  mutate(COUNT = round(FV_WEIGHT / 0.240)) |> 
  mutate(WEIGHT_MT = round(FV_WEIGHT / 1000), WEIGHT_KG = round(FV_WEIGHT, -3), COUNT = round(COUNT, -3)) |> 
  mutate(YEAR = as.factor(YEAR)) |> 
  ungroup() |> # Ungroup here so that the refactoring of the RIVERNAME works
  mutate(COUNTY = fct_reorder(COUNTY, WEIGHT_MT, .fun = max, .desc = TRUE)) -> ys_catch_county

ys_catch_county

ys_catch_county |> 
  ggplot(aes(YEAR, WEIGHT_MT, fill = COUNTY)) +
  geom_col(position = position_dodge(), colour = "#2F4F4F") +
  labs(
    title = "Annual reported gaspereau landings for Yarmouth / Shelburne",
    y = "Landings (mt)"
    ) +
  scale_y_continuous(
    limits = c(0, 8e2),
    breaks = seq(0, 8e2, 100),
    labels = scales::label_comma()
  ) +
  theme_bw() +
  theme(
    legend.position = "right",
    plot.title = element_text(size = 16),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    panel.grid.minor = element_blank()
    ) +
  xlab("")

ggsave("landings_ys.png", plot = last_plot(), width = 9, height = 6, units = "in", dpi = 300)

```
