---
title: "`r paste('Annual gaspereau escapement assessment for Tusket River for', format(Sys.Date(), '%Y'))`"
output: html_document
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,    # Hide code
  results = 'hide',# Hide results
  warning = FALSE, # Hide warnings
  message = FALSE  # Hide messages
)
```

```{r library_etc}
# Library
library(tidyverse)
library(padr) # for dates
library(ROracle)
library(imputeTS)

# Vars
current_year <- 2026

# Source necessary functions from the Alosa functions
source("~/git/ALOSA.functions/functions/sourcery.R")
sourcery()

channel <- dbConnect(
  DBI::dbDriver("Oracle"),
  username = "GASPEREA",
  password = "_REMOVED",
  "PTRAN",
  believeNRows = FALSE
  )

# Useful to remember for plots
#RColorBrewer::display.brewer.all()

```

```{r theme}
# You can set up a theme here if you want so that plotting is easier
# You can edit your theme here and then just add it in like an object
# custom_theme <- theme_minimal(base_family = "sans") +
#   theme(
#     plot.background = element_rect(fill = "#1e1e1e", color = NA),
#     panel.background = element_rect(fill = "#1e1e1e", color = NA),
#     panel.grid.major = element_line(color = "#333333"),
#     panel.grid.minor = element_blank(),
#     axis.text = element_text(color = "white"),
#     axis.title = element_text(color = "white"),
#     plot.title = element_text(color = "white", size = 16, face = "bold"),
#     legend.background = element_rect(fill = "#1e1e1e"),
#     legend.key = element_rect(fill = "#1e1e1e"),
#     legend.text = element_text(color = "white"),
#     legend.title = element_text(color = "white")
#   )
```

# Pre-season

Only run this at the beginning of the season to generate the files we will be
filling out with data throughout the season.

```{r create_data_sheets}

# Creates blank data sheets for counts and biodata for Lake Vaughan
# blank.datasheets(
#   seed = 118,
#   startmonth = 4,
#   endmonth = 6,
#   startday = 1,
#   rivername = "Vaughan",
#   year = 2026,
#   recordtime = TRUE,
#   speciesID = TRUE,
#   strata = 6,
#   samplesperstrata = 4
# )
# 
# # Creates blank data sheets for counts and biodata for Powerhouse
# blank.datasheets(
#   seed = 118,
#   startmonth = 4,
#   endmonth = 6,
#   startday = 15,
#   rivername = "Powerhouse",
#   year = 2026,
#   recordtime = TRUE,
#   speciesID = TRUE,
#   strata = 6,
#   samplesperstrata = 4
# )

```

Generate a list of files based on the data sheets created above so we can use it
to upload the correct videos to the shared drive.

```{r}
#file_names_csv <- read_csv("C:/Users/graylo/OneDrive - DFO-MPO/Alosa/Counts/2026/Vaughan 2026 count data.csv")
#file_names_csv <- read_csv("C:/Users/graylo/OneDrive - DFO-MPO/Alosa/Counts/2026/Powerhouse 2026 count data.csv")

# Add leading zeros to all values in the month column, days < 10, and times so
# that all times have four characters

# file_names_csv |>
#   mutate(mon = paste0("0", mon)) |>
#   mutate(day = as.character(day)) |>
#   mutate(day = if_else(day %in% c(1:9), paste0("0", day), day)) |>
#   # This is a function for string formatting from C. It is saying that the value
#   # is an integer (the %d) and it should be 4 characters wide with leading zeros
#   # if necessary (04)
#   mutate(time = sprintf("%04d", time)) |>
#   # mutate(new_date_time = paste0("Vaughan-", day, mon, year, "-", time, ".h264")) |>
#   mutate(new_date_time = paste0("POWERHOUSE-", day, mon, year, "-", time, ".h264")) |>
#   pull(new_date_time) -> file_names
# 
# #writeLines(file_names, "C:/Users/graylo/OneDrive - DFO-MPO/Alosa/Counts/2026/Vaughan 2026 Count filenames for video uploader script.txt")
#writeLines(file_names, "C:/Users/graylo/OneDrive - DFO-MPO/Alosa/Counts/2026/Powerhouse 2026 Count filenames for video uploader script.txt")

```


# In-season

## VD daily escapement
```{r escapement_vd}
# Locate your CSV for counts
#path_to_counts_CSV <- "C:/Users/graylo/Documents/Vaughan 2025 count data.csv"
path_to_counts_CSV <- "C:/Users/graylo/OneDrive - DFO-MPO/Alosa/Counts/2026/Vaughan 2026 count data.csv"

# Run escapement script
# Note: arguments year, site, and channel don't matter when database = FALSE
escapement_vd <- onespecies.river.escapement(
  path_to_counts_CSV,
  fixtime = TRUE,
  downstream.migration = FALSE,
  database = FALSE,
  year = current_year,
  site = 2,
  channel = channel
)

# We want to add where the data are from
escapement_vd$location <- "Lake Vaughan"

escapement_vd  |> 
  mutate(across(everything(), ~ ifelse(is.nan(.), 0, .))) |> 
  mutate(across(everything(), ~ ifelse(is.na(.), 0, .))) |> 
  mutate(date = make_date(current_year, mon, day)) -> escapement_vd

#in_season_vd_escapement_plot <-
escapement_vd |> 
  ggplot(aes(date, total)) +
  geom_ribbon(
    aes(ymin = clow, ymax = chigh),
    fill = "#41B7C4",
    colour = "#4BBAC3",
    alpha = 0.2
  ) +
  geom_path(
    colour = "#41B7C4",
    alpha = 0.9,
    linewidth = 1
    ) +
  theme_bw() +
  # geom_smooth(
  #   method = "loess",
  #   # Adjust the span to control smoothness (lower = more flexible)
  #   span = 0.3,
  #   se = FALSE,
  #   colour = "#F6AC50",
  #   linewidth = 1
  #   ) +
  # You can highlight regions using something like this
  #annotate("rect", xmin = as.Date("2025-04-25"), xmax = as.Date("2025-05-10"), ymin = 0, ymax = Inf, alpha = 0.1, fill = "orange")+
  labs(
    title = paste("Gaspereau escapement at the Lake Vaughan fish ladder in", current_year),
    x = "Date",
    y = "Fish / day"
  ) +
  # scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  # scale_y_continuous(
  #   limits = c(0, max(escapement_vd$chigh, na.rm = TRUE)),
  #   breaks = seq(0, max(escapement_vd$chigh, na.rm = TRUE), by = 1e4),
  #   labels = scales::comma
  # ) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      vjust = 1,
      hjust = 1
      ),
    axis.title.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "grey95")
    )

# Can use this for an interactive plot
#ggplotly(in_season_vd_escapement_plot)

# Save as a png for sharing
#ggsave("C:/Users/graylo/Documents/in_season_escapement_vd.png", plot = in_season_vd_escapement_plot, width = 6, height = 4, dpi = 300)
```

## PH daily escapement
```{r escapement_ph}
# Locate your CSV for counts
path_to_counts_CSV <- "C:/Users/graylo/OneDrive - DFO-MPO/Alosa/Counts/2026/Powerhouse 2026 Count data.csv"

# Run escapement script
# Note: arguments year, site, and channel don't matter when database = FALSE
escapement_ph <- onespecies.river.escapement(
  path_to_counts_CSV,
  fixtime = TRUE,
  downstream.migration = FALSE,
  database = FALSE,
  year = current_year,
  site = 14,
  channel = channel
)

# We want to add where the data are from
escapement_ph$location <- "Powerhouse"

escapement_ph |> 
  mutate(across(everything(), ~ ifelse(is.nan(.), 0, .))) |> 
  mutate(across(everything(), ~ ifelse(is.na(.), 0, .))) |> 
  mutate(date = make_date(current_year, mon, day)) -> escapement_ph

escapement_ph |> 
  ggplot(aes(date, total)) +
  geom_path(
    colour = "#7ECDBB",
    alpha = 0.9,
    linewidth = 1.25
    ) +
  geom_ribbon(
    aes(ymin = clow, ymax = chigh),
    fill = "#7ECDBB",
    colour = "grey90",
    alpha = 0.2
  ) +
  theme_bw() +
  labs(
    title = paste("Gaspereau escapement for Powerhouse for", current_year),
    x = "Date",
    y = "Fish / day"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 days") +
  scale_y_continuous(
    limits = c(0, max(escapement_ph$chigh, na.rm = TRUE)),
    breaks = seq(0, max(escapement_ph$chigh, na.rm = TRUE), by = 1000),
    labels = scales::comma
  ) +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    axis.title.x = element_blank(),
    panel.grid.minor = element_blank(),
    #legend.background = element_blank(),
    plot.background = element_rect(fill = "grey95")
  )

# plot_path <- paste0(output_folder, "/in_season_ph_escapement_plot.png")
#     
#     ggsave(
#       plot_path,
#       plot = in_season_ph_escapement_plot,
#       width = 6,
#       height = 4,
#       dpi = 300
#     )
```

## PH cumulative
```{r}
escapement_ph |> 
  arrange(date) |> 
  mutate(cumulative_total = cumsum(total)) |> 
  ggplot(aes(date, cumulative_total)) +
  geom_path(alpha = 0.8, linewidth = 1, na.rm = TRUE)+
  theme_bw() +
  labs(
    title = "Cumulative escapement of gaspereau at Powerhouse",
    x = "Date",
    y = "Total fish escaped"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  scale_y_continuous(
    limits = c(0, 1e4),
    breaks = seq(0, 1e4, 1e3),
    labels = scales::comma
  ) +
  theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.background = element_blank(),
    plot.background = element_rect(fill = "grey95"),
    legend.key = element_blank()
  ) +
  guides(colour = guide_legend(override.aes = list(linewidth = 2)))
```

## VD multi-year
```{r compare_vd}
# Set up a function to haul data from the db for VD
get_VD_multiple_years <- function() {
  
  counts_df <- data.frame()
  
  # You can change how many years you want here
  get_years <- seq(as.integer(current_year) - 5, as.integer(current_year) - 1)
  
  for (i in get_years) {
    
    tryCatch({
      
      year_data <- onespecies.river.escapement(
        count.data = "dummy_file_name",
        fixtime = TRUE,
        downstream.migration = TRUE,
        database = TRUE,
        year = i,
        site = 2,
        channel = channel
      )
      
      year_data$year <- i
      
      counts_df <- rbind(counts_df, year_data)
      
    }, error = function(e) {
      
      return(counts_df)
      
    })
    
  }
  
  return(counts_df)
  
}

multi_year_counts <- get_VD_multiple_years()

# Change some columns so we can merge
escapement_vd |> 
  select(-location, -date) |> 
  mutate(year = current_year) -> vd_for_merge

multi_year_counts <- rbind(multi_year_counts, vd_for_merge)

# Make sure to filter before creating dates because this screws them up ¯\_(ツ)_/¯
multi_year_counts <- multi_year_counts |> 
  mutate(across(everything(), ~ ifelse(is.nan(.), 0, .))) |> 
  mutate(across(everything(), ~ ifelse(is.na(.), 0, .))) |> 
  # You make these all have the same year in their date so they plot nicely
  mutate(date = make_date(as.character(current_year), mon, day)) |> 
  mutate(year = as.character(year)) -> multi_year_counts

multi_year_counts |> 
  filter(year > 2020) -> temp_df

up_to_date <- last(temp_df$date)

temp_df |> 
  group_by(year) |> 
  ggplot(aes(date, total)) +
  geom_path(
    aes(colour = year),
    alpha = if_else(temp_df$year != 2026, 0.4, 0.8),
    linewidth = if_else(temp_df$year != 2026, 1.3, 2),
    na.rm = TRUE
  ) +
  geom_point(
    aes(
      x = temp_df |> filter(year == 2026) |> last() |> pull(date),
      y = temp_df |> filter(year == 2026) |> last() |> pull(total),
      fill = year
      ),
    size = 3,
    pch = 21,
    alpha = 0.1
    ) +
  theme_bw() +
  labs(
    title = paste0("Gaspereau escapement at Lake Vaughan Dam as of ", up_to_date),
    x = "Date",
    y = "Fish / day",
    colour = "Year",
    fill = "Year",
    linetype = "Location"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  scale_y_continuous(
    limits = c(0, max(temp_df$total)),
    breaks = seq(0, max(temp_df$total), by = 10000),
    labels = scales::comma
  ) + 
  theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.background = element_blank(),
    plot.background = element_rect(fill = "grey95"),
    legend.key = element_blank()
  ) +
  guides(
    colour = guide_legend(override.aes = list(linewidth = 2)),
    fill = guide_legend(override.aes = list(size = 0))
    )

# plot_path <-
#   paste0(output_folder, "/in_season_multi_year_counts_plot.png")
# 
# ggsave(
#   plot_path,
#   plot = in_season_multi_year_counts_plot,
#   width = 6,
#   height = 4,
#   dpi = 300
# )
```

## VD average
```{r}
temp_df |> 
  mutate(flag = if_else(year < 2026, "2021 - 2025 Average", "2026")) |> 
  group_by(flag, date) |> 
  summarise(total = mean(total, na.rm = TRUE)) -> avg_esc_df

avg_esc_df |> 
  ggplot(aes(date, total)) +
  geom_path(aes(colour = flag), size = 1.5, alpha = 0.8) +
  theme_bw() +
  labs(
    title = "Average Gaspereau escapement at Lake Vaughan Dam",
    x = "Date",
    y = "Fish / day",
    colour = "",
    fill = "",
    linetype = "Location"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  scale_y_continuous(
    limits = c(0, max(avg_esc_df$total)),
    breaks = seq(0, max(avg_esc_df$total), by = 10000),
    labels = scales::comma
  ) + 
  theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.background = element_blank(),
    plot.background = element_rect(fill = "grey95"),
    legend.key = element_blank()
  )
```


## VD multi-year cumulative
```{r}
# Multi-year cumulative counts

multi_year_cum <- multi_year_counts |> 
  group_by(year) |> 
  arrange(date) |> 
  mutate(cumulative_total = cumsum(total))

# multi_year_cum |> 
#   group_by(year)|> 
#   ggplot(aes(date, cumulative_total)) +
#   geom_path(
#     aes(colour = year),
#     alpha = 0.8,
#     linewidth = 2,
#     na.rm = TRUE
#   )+
#   #scale_colour_viridis_d(direction = -1)+
#   theme_bw() +
#   labs(
#     title = "Cumulative escapement of gaspereau at Lake Vaughan Dam",
#     subtitle = "LRP indicated by dashed line",
#     x = "Date",
#     y = "Total fish escaped",
#     colour = "Year",
#     fill = "Year",
#     linetype = "Location"
#   ) +
#   scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
#   scale_y_continuous(
#     limits = c(0, 2.5e6),
#     breaks = seq(0, 2.5e6, 0.2e6),
#     labels = scales::comma
#   ) + 
#   geom_hline(yintercept = 2.086e6, linetype = "dashed")+
#   theme(
#     panel.grid.minor = element_blank(),
#     axis.title = element_text(face = "bold"),
#     axis.title.x = element_blank(),
#     axis.text.x = element_text(angle = 45, hjust = 1),
#     legend.background = element_blank(),
#     plot.background = element_rect(fill = "grey95"),
#     legend.key = element_blank()
#   ) +
#   guides(colour = guide_legend(override.aes = list(linewidth = 2)))

# Same plot but with the USR
multi_year_cum |> 
  group_by(year)|> 
  ggplot(aes(date, cumulative_total)) +
  geom_path(
    aes(colour = year),
    alpha = 0.8,
    linewidth = 2,
    na.rm = TRUE
  )+
  #scale_colour_viridis_d(direction = -1)+
  theme_bw() +
  labs(
    title = "Cumulative escapement of gaspereau at Lake Vaughan Dam",
    subtitle = "USR = dashed line, LRP = solid line",
    x = "Date",
    y = "Total fish escaped",
    colour = "Year",
    fill = "Year",
    linetype = "Location"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  scale_y_continuous(
    limits = c(0, 3.4e6),
    breaks = seq(0, 3.4e6, 0.25e6),
    labels = scales::comma
  ) + 
  geom_hline(yintercept = 2.086e6, linetype = "solid")+
  geom_hline(yintercept = 3.1e6, linetype = "dashed")+
  theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.background = element_blank(),
    plot.background = element_rect(fill = "grey95"),
    legend.key = element_blank()
  ) +
  guides(colour = guide_legend(override.aes = list(linewidth = 2)))
```

```{r}
# Alternative plot with years as continuous variable to show how newer years are
# getting lower (lighter colours)

multi_year_cum <- multi_year_counts |> 
  group_by(year) |> 
  arrange(year, date) |> 
  mutate(cumulative_total = cumsum(total), year = as.integer(year))

multi_year_cum |> 
  ggplot(aes(date, cumulative_total)) +
  geom_path(
    aes(colour = year, group = year),
    alpha = 0.8,
    linewidth = 2,
    na.rm = TRUE
  )+
  theme_bw() +
  labs(
    title = "Cumulative escapement of gaspereau at Lake Vaughan Dam",
    x = "Date",
    y = "Total fish escaped",
    colour = "Year",
    fill = "Year",
    linetype = "Location"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  scale_y_continuous(
    limits = c(0, 2.5e6),
    breaks = seq(0, 2.5e6, 0.2e6),
    labels = scales::comma
  ) + 
  theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.background = element_blank(),
    plot.background = element_rect(fill = "grey95"),
    legend.key = element_blank()
  ) +
  guides(colour = guide_legend(override.aes = list(linewidth = 2)))
```

## VD cumulative with CI
```{r}

escapement_vd |> 
  arrange(date) |> 
  mutate(
    ctotal = cumsum(total),
    cclow = cumsum(clow),
    cchigh = cumsum(chigh)
    ) |>  
  ggplot(aes(date, ctotal)) +
  geom_ribbon(aes(ymin = cclow, ymax = cchigh), alpha = 0.4) +
  geom_path(linewidth = 1, na.rm = TRUE) +
  geom_path(aes(date, cchigh), linewidth = 0.5) +
  geom_path(aes(date, cclow), linewidth = 0.5) +
  theme_bw() +
  labs(
    title = "Cumulative escapement of gaspereau at Lake Vaughan Dam in 2026",
    x = "Date",
    y = "Total fish escaped"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 day") +
  # scale_y_continuous(
  #   limits = c(0, 3e5),
  #   breaks = seq(0, 3e5, 2e4),
  #   labels = scales::comma
  # ) +
  theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.background = element_blank(),
    plot.background = element_rect(fill = "grey95"),
    legend.key = element_blank()
  ) +
  guides(colour = guide_legend(override.aes = list(linewidth = 2)))
```


## PH multi-year
```{r multi_year_ph}
# 2022, 2023, 2024
ph_counts_2022 <- onespecies.partial.river.escapement(
  "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2022/Data Sheets/Counts/Powerhouse Count Sheet Cleaned 2022.csv",
  fixtime = FALSE,
  database = FALSE,
  year = 2022,
  site = 14,
  channel = channel
  )
ph_counts_2022$year <- as.character(2022)
ph_counts_2022$location <- "Powerhouse"

ph_counts_2023 <- onespecies.river.escapement(
  "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2023/Data Sheets/Powerhouse 2023 count data1.csv",
  fixtime = FALSE,
  downstream.migration = FALSE,
  database = FALSE,
  year = 2023,
  site = 14,
  channel = channel
  )
ph_counts_2023$year <- as.character(2023)
ph_counts_2023$location <- "Powerhouse"

ph_counts_2024 <- onespecies.river.escapement(
  filename = NULL,
  fixtime = TRUE,
  downstream.migration = FALSE,
  database = TRUE,
  year = 2024,
  site = 14,
  channel = channel
)
ph_counts_2024$location <- "Powerhouse"
ph_counts_2024$year <- 2024

# 2025
ph_counts_2025 <- onespecies.river.escapement(
  "C:/Users/graylo/Documents/Powerhouse 2025 count data.csv",
  fixtime = TRUE,
  downstream.migration = FALSE,
  database = FALSE,
  year = 2025,
  site = 14,
  channel = channel
)
ph_counts_2025$location <- "Powerhouse"
ph_counts_2025$year <- 2025

rbind(ph_counts_2025, ph_counts_2024, ph_counts_2023, ph_counts_2022) -> ph_counts

# Clean up all PH data
ph_counts |> 
  mutate(across(everything(), ~ ifelse(is.nan(.), 0, .))) |> 
  mutate(across(everything(), ~ ifelse(is.na(.), 0, .))) |> 
  select(-location) |>
  # Give same year so they plot nicely
  mutate(date = make_date(current_year, mon, day)) |>
  mutate(year = as.character(year)) -> ph_counts

# Escapement estimates
ph_counts |> 
  group_by(year) |>  
  ggplot(aes(date, total, fill = year)) +
  geom_path(
    aes(colour = year),
    alpha = 0.8,
    linewidth = 1,
    na.rm = TRUE
  )+
  geom_ribbon(
    aes(ymin = clow, ymax = chigh),
    # fill = "#7ECDBB",
    # colour = "grey90",
    alpha = 0.2
  ) +
  theme_bw() +
  labs(
    title = "Gaspereau escapement at the Powerhouse",
    x = "Date",
    y = "Fish / day",
    colour = "Year",
    fill = "Year"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  scale_y_continuous(
    limits = c(0, max(ph_counts$chigh, na.rm = TRUE)),
    breaks = seq(0, max(ph_counts$chigh, na.rm = TRUE), by = 2000),
    labels = scales::comma
  ) +
  theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.background = element_rect(fill = "grey95"),
    legend.background = element_blank(),
    legend.key = element_blank()
  )
```

## PH multi-year cumulative

```{r multi_year_cumulative}
# Multi-year cumulative for PH
ph_counts |> 
  group_by(year) |> 
  arrange(date) |> 
  mutate(cumulative_total = cumsum(total)) -> ph_cumulative

ph_cumulative |> 
  group_by(year) |> 
  ggplot(aes(date, cumulative_total)) +
  geom_path(
    aes(colour = year),
    alpha = 0.8,
    linewidth = 2,
    na.rm = TRUE
  )+
  theme_bw() +
  labs(
    title = "Cumulative escapement of gaspereau at the Powerhouse",
    x = "Date",
    y = "Total fish escaped",
    colour = "Year",
    fill = "Year",
    linetype = "Location"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  scale_y_continuous(
    limits = c(0, max(ph_cumulative$cumulative_total, na.rm = TRUE)),
    breaks = seq(0, max(ph_cumulative$cumulative_total, na.rm = TRUE), 1e4),
    labels = scales::comma
  ) +
  theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.background = element_blank(),
    plot.background = element_rect(fill = "grey95"),
    legend.key = element_blank()
  ) +
  guides(colour = guide_legend(override.aes = list(linewidth = 2)))
```


# Post-season

The analyses here can only be done after uploading the data from the year to the
database. Read the `Writing Annual Data to Oracle.R` script in the `Alosa.functions`
folder to see the protocol for uploading data. There are methods in that script
to prevent errors, check the data, etc.

## Uploading bio data
```{r}
# Use the checkers function to check the biocharacteristics data for outliers
# and mistakes
checkers(filename = "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2025/data/Vaughan 2025 biocharacteristics data.csv")

# Once you go through the checkers function, you can add the accessory data
# (i.e. the additional species ID checks we do for bluebacks) to the biodata.
# The accessory data have to be done a certain way for this to work. The data 
# need to be categorized as BB identified and all fish identified. This may
# be different than you record the data throughout the season, so make sure that
# your data are represented properly:
# Column names: mon, day, BB.sampled, all.sampled, BB.ID, all.ID
accessory <- read_csv(file = "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2025/data/vaughan-accessory-data-version-2.csv")

# WARNING: this writes to the "biofilename" file, it appends the data to the
# end of the file. If you run this more than one it will keep appending the
# data to the file. NOTE: make sure the order of the columns for the date in 
# the biocharacteristics CSV are as follows: day, mon, year
speciesID.expand(
  data = accessory,
  # The biocharacteristics data file path goes here
  biofilename = "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2025/data/Vaughan 2025 biocharacteristics data.csv",
  rivername = "Tusket",
  year = 2025
)

# Here are the new biodata
biodata <- read_csv(file = "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2025/data/Vaughan 2025 biocharacteristics data.csv")

# Look through the "Writing Annual Data to Oracle.R" script for more details
oracle.bio <- format.BIODATA.onesite("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2025/data/Vaughan 2025 biocharacteristics data.csv")

# You can get someone else to double check the data just to be safe
#saveRDS(oracle.bio, file = "biodata_for_checking.rds")

# Once you are happy with these data, upload to the GASPERA database.
# NOTE: make sure overwrite = FALSE
# dbWriteTable(
#   conn = channel,
#   schema = "GASPEREA",
#   value = oracle.bio,
#   name = "ALOSA_FISH_BIO_DATA",
#   date = TRUE,
#   row.names = FALSE,
#   overwrite = FALSE, ## NEVER CHANGE TO TRUE
#   append = TRUE
# )

```

## Uploading count data
```{r}
# Go through the count data to double check things. Things to look for:
# - Check for really big or really small variances (maybe a keystroke error)
# - Gaps
# - Did we miss any counts, do we have the footage to redo it? 
# - Large number of downstream movement.

oracle.count <- format.COUNTDATA.onesite("C:/Users/graylo/Documents/Vaughan 2025 count data.csv")

# If you are happy with it, upload to the GASPERA database
# NOTE: make sure overwrite = FALSE
# dbWriteTable(
#   conn = channel,
#   schema = "GASPEREA",
#   value = oracle.count,
#   name = "ALOSA_VIDEO_COUNT_DATA",
#   date = TRUE,
#   row.names = FALSE,
#   ## NEVER CHANGE TO TRUE
#   overwrite = FALSE,
#   append = TRUE
# )
```

## Selecting scales
```{r}
# We use the ageing.selection.v1.1 script to select which scales we choose to
# process for ageing. It uses the daily.summary object from the
# onespecies.river.escapement script.
temp <- onespecies.river.escapement(
  "does_not_matter",
  fixtime = TRUE,
  downstream.migration = FALSE,
  database = TRUE,
  year = current_year,
  site = 2,
  channel = channel
)

# Biodata from CSV - this doesn't work currently with biodata taken from the db
biodata <- read.csv("R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2025/data/Vaughan 2025 biocharacteristics data.csv")

biodata <- standardise_column_names(biodata)

# Calculate missing days
missing.days(biodata)

# This selects the scales to age
# Note: we don't put anything in the mergedays argument because this is only
# needed if weekly = FALSE
ageing.selection.test(
  countdata = temp,
  biodata = biodata,
  # This weights samples by the week and is used when weekends were not sampled
  # We use this at Tusket since we do not sample on weekends because we are
  # cool and based
  weekly = TRUE,
  year = current_year,
  missingdays = missing_days
  )

# This will output a csv file to the working directory

# Check to see if any day has almost all of the fish being sampled for aging
scales_sampled_per_time_unit |> 
  group_by(weekofyear) |>
  summarise(
    count_to_be_aged = sum(!is.na(to.be.aged)),
    n_sampled = first(n.sampled),
    .groups = "drop"
  )
```

## Species proportions
```{r species}
# Get the counts for the year for Lake Vaughan
counts <- onespecies.river.escapement(
  "lololololol",
  fixtime = TRUE,
  downstream.migration = FALSE,
  database = TRUE,
  year = current_year,
  site = 2,
  channel = channel
)

counts <- counts %>%
    mutate(across(everything(), ~ ifelse(is.nan(.), 0, .))) %>%
    mutate(across(everything(), ~ ifelse(is.na(.), 0, .)))

counts <- counts |> mutate(date = make_date(current_year, mon, day))

counts$year <- current_year

# This calculates the proportions of alewives and BBs in the escapement estimates
# based on the biodata where we took information on species proportions.
proportions <- split_species(
    year = current_year,
    siteID = 2,
    channel = channel
  )

counts_species <- counts |>
      mutate(date = make_date(year, mon, day)) |>
      ungroup() |>
      select(date, total, clow, chigh) |>
      inner_join(proportions, by = "date") |>
      mutate(B = total * BBprop, A = total - B) |>
      select(date, A, B) |>
      pivot_longer(cols = c(A, B),
                   names_to = "species",
                   values_to = "total") |>
      mutate(species = str_replace(species, "A", "Alewives")) |>
      mutate(species = str_replace(species, "B", "Bluebacks")) |>
      ungroup()

# This plot shows total counts of each species - this is not very useful
# because there are so few BBs
# counts_species |>
#   group_by(species) |>
#   ggplot(aes(date, total)) +
#   geom_path(aes(colour = species), size = 1) +
#   geom_point(aes(fill = species), shape = 21, colour = "black", size = 2)+
#   theme_bw() +
#   labs(
#     title = paste0(
#       "Escapement estimates for gaspereau by species for ",
#       current_year
#     ),
#     y = "Fish / day",
#     colour = "Species",
#     fill = "Species"
#   ) +
#   scale_x_date(date_labels = "%b %d", date_breaks = "3 days") +
#   scale_y_continuous(
#     limits = c(0, max(counts_species$total, na.rm = TRUE)),
#     breaks = seq(0, max(counts_species$total, na.rm = TRUE), by = 10000),
#     labels = scales::comma
#   ) +
#   theme(
#     axis.text.x = element_text(
#       angle = 45,
#       vjust = 1,
#       hjust = 1
#       ),
#     plot.background = element_rect(fill = "grey95"),
#     legend.background = element_blank(),
#     legend.key = element_blank(),
#     axis.title.x = element_blank(),
#     panel.grid.minor = element_blank()
#   )

# Here is a plot looking at percentages of species over time
# Get the proportions
counts_species |> 
  group_by(date) |> 
  mutate(percent = total / sum(total) * 100) |> 
  ungroup() |> 
  filter(!is.nan(percent)) -> species_props

species_props |> 
group_by(species) |>
  filter(date > "2025-05-26") |> 
  ggplot(aes(date, percent)) +
  geom_path(aes(colour = species), size = 1) +
  geom_point(aes(fill = species), shape = 21, colour = "black", size = 2)+
  theme_bw() +
  labs(
    title = paste0(
      "Proportions of escaped fish at Lake Vaughan by species for ",
      current_year
    ),
    y = "Percent",
    colour = "Species",
    fill = "Species"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 day") +
  scale_y_continuous(
    limits = c(0, 100),
    breaks = seq(0, 100, 10),
    labels = scales::percent_format(scale = 1)
  ) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      vjust = 1,
      hjust = 1
      ),
    plot.background = element_rect(fill = "grey95"),
    legend.background = element_blank(),
    legend.key = element_blank(),
    axis.title.x = element_blank(),
    panel.grid.minor = element_blank()
  )

```

## Species split array
```{r}
# We can also use the twospecies.river.escapement function to get the species
# splits, but this has the added benefit of calculating CIs. We use the 
# proportions df we made above in the function
species_df <- twospecies.river.escapement(
  filename = "",
  year = current_year,
  site = 2,
  channel = channel,
  species.split = proportions
)

# Alewives
species_df[[1]]

# Bluebacks
species_df[[2]] -> BB

# Deal with the NaNs so that the plot is good
BB %>%
    mutate(across(everything(), ~ ifelse(is.nan(.), 0, .))) %>%
    mutate(across(everything(), ~ ifelse(is.na(.), 0, .))) -> BB

BB |>
  mutate(date = make_date(year = current_year, month = mon, day = day)) |> 
  ggplot(aes(date, total)) +
  geom_path(
    colour = "#7ECDBB",
    alpha = 0.9,
    linewidth = 1.25
    ) +
  geom_ribbon(
    aes(ymin = clow, ymax = chigh),
    fill = "#7ECDBB",
    colour = "grey90",
    alpha = 0.2
  ) +
  theme_bw() +
  labs(
    title = paste0("Blueback escapement at the Lake Vaughan fish ladder in ", current_year),
    x = "Date",
    y = "Fish / day"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 day") +
  scale_y_continuous(
    limits = c(0, max(BB$chigh)),
    breaks = seq(0, max(BB$chigh), by = 1e2),
    labels = scales::comma
  ) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      vjust = 1,
      hjust = 1
      ),
    axis.title.x = element_blank(),
    panel.grid.minor = element_blank(),
    #legend.background = element_blank(),
    plot.background = element_rect(fill = "grey95")
    )

# Cumulative plot of BBs
BB_cumulative <- BB |> mutate(cumulative_total = cumsum(total))

BB_cumulative |> 
  mutate(date = make_date(year = current_year, month = mon, day = day)) |> 
  ggplot(aes(date, cumulative_total)) +
  geom_path(colour = "purple", alpha = 0.8, linewidth = 1.75, na.rm = TRUE)+
  theme_bw() +
  labs(
    title = paste("Cumulative escapement of Blueback Herring at Lake Vaughan Dam in", current_year),
    x = "Date",
    y = "Total fish escaped",
    linetype = "Location"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 day") +
  scale_y_continuous(
    limits = c(0, max(BB_cumulative$cumulative_total)),
    breaks = seq(0, max(BB_cumulative$cumulative_total), 2e2),
    labels = scales::comma
  ) +
  theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.background = element_blank(),
    plot.background = element_rect(fill = "grey95"),
    legend.key = element_blank()
  ) +
  guides(colour = guide_legend(override.aes = list(linewidth = 2)))
```


## Species catch

These numbers are required for the `Tusket River Annual Summary` document that is
made every year. They are estimates of the number of bluebacks caught in the Tusket.
The numbers here are used to fill in data in `river summaries all years.csv` that
from which `Tusket River Annual Summary Document_Script.Rmd` pulls its data when writing
the document.

```{r blueback_catch_inspections}
inspection <- readxl::read_xlsx(
  "R:/Science/Population Ecology Division/DFD/Alosa/Locations/Tusket River/Tusket 2022/Data Sheets/2022 Yarmouth County Catch Inspection.xlsx",
  sheet = "Sheet1"
  )

inspection |>
  mutate(date = make_date(year = year, month = mon, day = day)) -> inspection

inspection |> 
  filter(
    date >= "2022-05-13" &
    date <= "2022-05-15" &
    method == "gill" &
    river == "Tusket") |> 
  summarize(BBprop = sum(n.blueback) / sum(n.alewife)) -> may_13_2022_prop

inspection |>
  filter(
    date >= "2022-05-20" &
    date <= "2022-05-22" &
    method == "gill" &
    river == "Tusket"
    ) |> 
  summarize(BBprop = sum(n.blueback) / sum(n.alewife)) -> may_20_2022_prop

```

```{r blueback_catch}
# This section will need to be edited for each year. The way this is currently
# set up requires you to identify which of the last weeks of May you want to
# calculate blueback catch for. This will be based on when bluebacks were
# detected. In the example below, we only used gill net data as there were no
# bluebacks found at the dipping stands when George and Mark sampled their
# catches in 2022.
load("R:/Science/Population Ecology Division/DFD/Alosa/MARFISSCI/Old pulls/marfis_pull.Rdata")
source("R:/Science/Population Ecology Division/DFD/Alosa/MARFISSCI/MARFIS.error.cleaner.R")
catch <- convert.KGS(catch)

# Total tusket catch
catch |>
  filter(YEAR == 2024 & RIVERNAME_CLEANED == "TUSKET") |> 
  # We divide by the conversion factor to convert mass to count
  summarise(total_catch = sum(KGS) / .240) -> tusket_total_catch

# I checked the logbooks for the Tusket licences from the filtered data above, and they all look good for 2024. There were a few small mistakes that I won't correct here.
catch |>
  filter(
    YEAR == 2024 &
    RIVERNAME_CLEANED == "TUSKET" &
    GEAR_DESCRIPTION == "GILL NET (SET OR FIXED)"
    ) -> tusket_gill

tusket_gill <- as_tibble(tusket_gill)

# Get landings per week for last two weeks for gill netters only
tusket_gill |>
  filter(
    FV_DATE_FISHED >= "2024-05-17" &
    FV_DATE_FISHED <= "2024-05-20") |>
  summarise(total_kg = sum(KGS)) -> week_1

tusket_gill |>
  filter(FV_DATE_FISHED >= "2024-05-24") |>
  summarise(total_kg = sum(KGS)) -> week_2

# Convert landings to individual fish using 0.240 kilograms per fish 
# (i.e. a fish has a mass of 240 g) and make tibble of all estimates
tibble(
  week1 = round(week_1$total_kg * may_13_2022_prop$BBprop / .240),
  week2 = round(week_2$total_kg * may_20_2022_prop$BBprop / .240)
  ) -> bb_catch

bb_catch |> 
  mutate(total_bb_catch = rowSums(across(everything()))) -> bb_catch

bb_catch$total_bb_catch -> bb_catch
```

```{r blueback_exploitation}
proportions$date <- ymd(proportions$date)

counts |> 
  mutate(date = make_date(year = year, month = mon, day = day)) -> counts

left_join(proportions, counts, by = "date") -> result

result |> 
  select(date, BBprop, total) |> 
  mutate(
    BB = BBprop * total,
    A = total - BB
  ) |>
  select(-BBprop, -total) |> 
  pivot_longer(
    cols = c(A, BB),
    names_to = "Species",
    values_to = "Total"
  ) -> result

result |> 
  group_by(Species) |> 
  summarise(sum = sum(Total)) -> bb_total

# Exploitation rate
bb_exploitation <- bb_catch / (bb_total$sum[2] + bb_catch)
bb_exploitation
```

```{r summary_table}
# Label counts from Lake Vaughan
counts$location <- "Lake Vaughan"
counts <- counts |> select(total, location)
total_count_vd <- sum(counts$total)

# PH counts
ph_counts <- onespecies.river.escapement(
  fixtime = FALSE,
  downstream.migration = FALSE,
  database = TRUE,
  year = 2024,
  site = 14,
  channel = channel
)
ph_counts$location <- "Powerhouse"
ph_counts <- ph_counts |> select(total, location)
total_count_ph <- sum(ph_counts$total)

# Combine counts from both sites
counts <- rbind(counts, ph_counts)
total_count <- sum(counts$total)

df <- tibble(
  tusket_escapement = total_count,
  VD_escapement = total_count_vd,
  PH_escapement = total_count_ph,
  VD_bb_escapement = bb_total$sum[2],
  tusket_catch = round(tusket_total_catch$total_catch),
  tusket_catch_bb = bb_catch
)

df <- df |> pivot_longer(cols = everything(), names_to = "variable", values_to = "count")
df <- df |> mutate(count = round(count))
df <- df |> mutate(count = scales::comma(count))

df
```

## Species GLM

Only need to do this if you don't check for BBs every day during the late run.
In 2024 we did not check on weekends, so we needed to use this to model the BBs.
In 2025 we did check on weekends, so we do not need it.
```{r GLM_approach}
# We often have days that we could not sample biodata for various reasons. We
# can use padr::pad to identify which dates are missing and then fill with NAs.
pp <- padr::pad(proportions, interval = "day")

# To prevent trouble down the road, we create a column to store the "date" as 
# an integer value. When date is used in the model, it really doesn't have to be
# a date, it just has to be an integer. As dates can cause headaches with many
# functions etc. it is safe to use an integer instead.
pp$day_int <- 1:nrow(pp)

# We use a Poisson GLM as we are dealing with count data. This model is imperfect
# and we know this, but for our current requirements, it is fairly useful. We
# wouldn't want to use this model to extrapolate very far out. We train the model
# on the complete data set.
model <- glm(BB ~ day_int, data = pp, offset = log(total_fish), family = "poisson")

# We set the total_fish to 1 so that we get proportions when we predict the new
# data. We usually try to get 100 fish sampled per day in the biodata, but that
# doesn't always happen, especially at the tails of the season.
out <- data.frame(day_int = 1:nrow(pp), total_fish = rep(1,nrow(pp)))
out.pred <- predict(model, newdata = out, type = "link", se.fit = T)

# This is the inverse of the link function i.e. the inverse of the log i.e exp
ginv <- model$family$linkinv

# We use the inverse of the link function to calculate the expected proportion.
# Note, this is the same as doing exp(out.pred[[1]])
out.pred <- ginv(out.pred[[1]])

# Create a new column for the new predicted proportions of BBs
pp$BBpropnew<-NA

# Populate the dataframe with the predicted values
for(i in 1:nrow(pp)) {
  if (is.na(pp$BBprop[i]) == T) {
    pp$BBpropnew[i] <- out.pred[i]
    pp$BBprop[i] <- out.pred[i]
  }
}

# We can flag which data were observed and which were modeled
pp$observation <- is.na(pp$total_fish)
pp <- pp |> mutate(observation = if_else(observation == TRUE, "Modeled", "Observed"))

# Check the plot
pp |> 
  ggplot(aes(date, BBprop))+
  geom_path(colour = "grey50", linewidth = 0.5, linetype = "dashed")+
  geom_point(aes(fill = observation), shape = 21, colour = "black", size = 2)+
  theme_bw()+
  labs(
    title = paste("Blueback herring proportions with values for missing dates from GLM for", current_year),
    y = "Proportion of fish",
    fill = "Estimate"
  ) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week") +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = .1),
    labels = scales::comma
  ) +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    plot.background = element_rect(fill = "grey95"),
    axis.title.x = element_blank(),
    legend.background = element_blank(),
    legend.key = element_blank()
  )

```
